### 2011-12-10 vm_snapshot_sim/related.tex

\section{Related Works}
Several approaches have been previously proposed to enable efficient deduplication in D2D backup.

DDFS\cite{bottleneck08} exploits chunk locality to achieve high-throughput perfect deduplication. 
It preserves locality by a Stream-Informed Segment Layout and exploits locality with Locality Preserved Cache. 
An in-memory Bloom Filter is also used to accelerate non-duplicate chunk identification.

Sparse Indexing\cite{sparseindex09} is an approximate deduplication technique designed for D2D backup. 
It divides data stream into variable-sized multiple kilobytes chunks, and construct multiple megabytes segments
using the same chunking technique, 
which are then sampled and mapped to a compact in-memory sparse index. 
Incoming segments are only deduplicated against several existing similar 
segments selected according to the sparse index.

Both DDFS and Sparse Indexing are designed for D2D backup workloads, 
and do not address the scalability issue in a distributed environment. 
A few scalable deduplication approaches have been proposed recently.

Extreme Binning\cite{extreme_binning09} is a scalable parallel deduplication approach 
that targets at non-traditional backup workloads that consist of low-locality individual files. 
It groups highly similar files into bins, and eliminates duplicate chunks inside each bin. 
Duplicate chunks are allowed to exist among different bins, resulting in approximate deduplication. 
By keeping only the primary index in memory, Extreme Binning can reduce the RAM requirement while 
maintaining a reasonably high throughput. However, their per-file based similarity detection 
is going to group all similar files into one node, which will break load balancing if some files are
huge and similar(e.g., virtual machine images).

MAD2\cite{mad210} is a scalable high-throughput exact duplication approach. 
MAD2 utilizes on-disk Hash Bucket Matrix to preserve fingerprint locality and 
integrates in-memory Dual Cache to capture and exploit locality. 
In addition, MAD2 employs Bloom Filter Array to efficiently identify unique 
incoming fingerprints and indicate where a duplicate may reside. 
By employing a DHT-based Load-Balance technique to distribute file recipes 
and chunk contents among multiple storage nodes in their backup sequences, 
MAD2 further enhances performance with a well balanced load. However, the
data locality does not exist for cloud storage because only changed chunks are expected
to be uploaded, and
the heavy usage of memory and CPU indicates such exact deduplication backup systems
need storage-exclusive servers with hardware replication support, which is not a general case 
for the cloud..

HYDRAstor\cite{hydrastor09}, a scalable secondary storage solution, 
constructs its backend using a grid of storage nodes built around a distributed hash table. 
The backend maintains large-scale variable-sized, content-addressed, immutable, 
and highly-resilient data blocks that are logically organized in a directed acyclic graph. 
Duplicate chunks are eliminated according to their hashes. 
HYDRAstor adopts an average chunk size of 64KB, among other constraints, 
to keep all the metadata in memory and avoid the duplicate-lookup disk bottleneck. 
This degrades the space efficiency of deduplication, and still requires huge amount of memory.

We believe a deduplication backend of cloud storage must have scalability and
 high availability built in mind, being able to run on low-cost non-proprietary machines.
  Unlike sloud, all above systems lack one or a few such properties. All exact deduplication approaches
are too costly, this is why we choose similarity based approach to trade deduplication
accuracy for speed.

Eariler deduplication systems mainly focus on improving storage space efficiency by eliminating 
duplicates at the file level, fixed-size block level, or variable-sized chunk level. 
EMC's Centera\cite{emc_centera} identify and eliminate duplicate data by comparing 
the hash of the whole file or fixed content. Venti\cite{venti02}, a block-level archival storage, 
removes redundant fixed-size data blocks by comparing their secure hashes. Pastiche\cite{pastiche02} 
utilizes chunk-level duplicate detection to construct a resource-saving peer-to-peer backup network. 
Deep Store\cite{deepstore05}, a large scale archival storage system, uses both variable-sized 
chunk-level deduplication and delta compression to save storage. Jumbo Store\cite{jumbo07} organizes 
variable-sized chunks into Hash-Based Directed Acyclic Graphs to save both storage and 
bandwidth while performing incremental upload and versioning for a utility rendering service.

Duplicate detection technique has also been used in bandwidth-saving synchronization protocols\cite{rsync} 
and low-bandwidth network file systems\cite{lbfs01}.