### 2012-01-30 vm_snapshot_sim/conclusion.tex

\section{Conclusion}
In this paper we present a new parallel deduplication technique for VM cloud.
Our technique utilizes the special data characteristic in VM cloud backup system
to accommodate very limited resources for data deduplication.
By storing the commonly used OS related data, and the hottest user generated data, 
we greatly reduce the cross-VM data duplication in VM snapshot backups. Experiments show
our solution can eliminate the majority of data duplication with a tiny fraction of
block hash index store in memory. It does not only saves valuable system resouces in
the VM cloud, but also makes deduplication much faster.