### 2012-01-30 vm_snapshot_sim/main.tex

\documentclass[finalversion]{usetex-v1}
%\usepackage[top=1in, bottom=1in, left=1.25in, right=1.25in]{geometry}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{epsfig}
%\usepackage{breakurl}
%% Define a new 'leo' style for the package that will use a smaller font.
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
\makeatother
%% Now actually use the newly defined style.
\urlstyle{leo}
%\linespread{1.2}
%\setlength{\parskip}{1ex}
\begin{document}
\title{Data Deduplication and Zipf-like Distribution in VM Cloud Storage}
\author{Wei Zhang, Hong Tang, Hao Jiang, Rachael Zeng, Xiaogang Li, Tao Yang}
\date{}
\maketitle

\input{abstract}
\input{intro} %introduce the problem, related solutions and our requirements
\input{cds} %data study, zipf, model of cds size and dedup efficiency
\input{approach} %our dedup process
\input{results}
%\input{related}
%\input{future}
\input{conclusion}
\bibliographystyle{abbrv}
\bibliography{dedup}
\end{document}