### 2012-02-29 vm_snapshot_sim/abstract.tex
Update abstract to a grammar corrected version.

\begin{abstract}
In a virtualized cloud computing environment,
frequent snapshot backups of virtual disks improve hosting reliability,
while storage demand of such operations is huge.
While a dirtybit-based technique can identify unmodified data
between versions, full deduplication with fingerprint comparison
can remove more redundant content but it requires more computing resources.
In this paper, we present a multi-level selective deduplication scheme which integrates inner-VM
and cross-VM duplicate elimination with minimal resource impact
to the existing cloud services. This scheme uses popular common
data to facilitate fingerprint comparison while reducing the cost,
and it strikes a balance between local and global deduplication to increase parallelism and improve
reliability. Experimental results show the proposed scheme can achieve a high deduplication
ratio while using a small amount of cloud resources.
\end{abstract}