### 2012-02-29 vm_snapshot_sim/2012background.tex

\section{Background and Related Work}
\label{sect:background}

In a VM cloud, several operations are provided for creating and managing snapshots and snapshot trees,
such as creating snapshots, reverting to any snapshot, and removing snapshots.
For VM snapshot backup, file-level semantics are normally not provided. 
Snapshot operations are taken place at the virtual device driver level, which means no fine-grained file system metadata can be used to determine the changed data. Only raw access information at disk block level are provided.
%Snapshots may be shared among VMs when a user runs the same binary image on different  VMs.
% This feature complicates the snapshot management, 
%but also require all snapshots be managed in a global namespace.

%While snapshot backup loads are heavy but computing and storage  resources available 
%can be limited to reduce the overall operation cost of a cloud. The whole VM cluster have petabytes of data and need to 
%be backed up at daily basis, if not more frequently. But at the same time snapshot tasks must not affect the normal VM operations
%in a cloud service, which means only a tiny slice of CPU and memory can be used for this purpose.

VM snapshots can be  backed up  incrementally by identifying file  blocks that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
%Even though the snapshots are saved incrementally,
%when deleting a snapshot, only the data not needed for any other snapshot is removed.
%So regardless of which prior snapshots have been deleted, all active snapshots will contain
%all the information needed to restore the virtual disk.
%One widely used snapshot strategy is copy-on-write (CoW).
%The earlier use of COW is to improve memory usage and reduce copying overhead in process 
%management~\cite{OSbook,Waldspurger2002} and it can be extended for snapshot 
%storage management~\cite{vmware_kb1015180}. 
%C. Waldspurger. Memory Resource Management in VMware ESX Server in Proceedings of the 5th Symposium 
%on Operating Systems Design and Implementation, 2002
%Upon VM image storage system receives a save snapshot request,
%it freezes the state of that image file, then all consequent write request will result in the write region being copied
%to an incremental snapshot data file. 
%Such a strategy exploits version difference between consecutive snapshots
%of the same image.
The main weakness
is that it does not reveal content redundancy among data blocks from different files or
different VMs.
%has several disadvantages:
%first, CoW may affect the general I/O performance due to defered data coping. 
%Second, CoW does not seperate backup data and runtime image data,
%which have distinct access requirements: runtime image data is directly used by the running VM, 
%thus need high throughput, and is very sensitive to latency,
%such data must be served with hig cost hardware, but backup data generally only need fair aggregate throughput, 
%is not sensitive to latency, thus can be stored in secondary storage devices.
%Finally, VM snapshots contain tremendous amount of data duplication, which is nearly impossible to tackle 
%if these two kinds of data are tightly coupled together.

Data deduplication techniques can eliminate redundancy globally among different files from different users. 
Backup systems have been developed to use content hash (finger prints) to identify duplicate 
content~\cite{venti02,Rhea2008}.
%,NGmiddleware2011}. 
Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper} 
use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
%Chunking divides a data stream into variable length chunks, it has been used to conserve bandwidth\cite{lbfs01}, 
%search and index documents in large repositories\cite{bhag07}, scalable storage systems\cite{hydrastor09}, 
%store and transfer large directory trees efficiently and with high reliability\cite{jumbo07}.
To chunk a file, starting from the first byte, its contents as seen through a fixed-sized (overlapping) 
sliding window are examined. At every position of the window, a fingerprint of its 
contents, $f$, is computed using hashing techniques such as Rabin fingerprints~\cite{rabin81}. 
When this fingerprint meets a certain criteria,
% such as $f mod D = r$ where $D$, the divisor, 
%and $r$ are predefined values; 
the position of the corresponding window defines the boundary of the chunk. 
This process is repeated until the entire file has been completely broken down into chunks. Next, 
a cryptographic hash or chunk ID of the chunk is computed using a technique such as MD5 or SHA.
After a file is chunked, the index containing the chunk IDs of backed up chunks 
is queried to determine duplicate chunks. New chunks are written to disk and the 
index is updated with their chunk IDs. A file recipe containing all the information 
required during reconstruction is generated. The index also contains some metadata 
about each chunk, such as its size and disk location.
%The compression rate obtained by deduplication depends on the inherent content overlapping degree in a dataset, 
%the average size of chunks, and the chunking method\cite{poli04}. 
%In general, smaller chunks yield better deduplication.

Data deduplication has been applied as  a post-processing operation for offline removal
of redundant content in the file system data management once backup data is written in a temporary stage area.
Dynamic inline deduplication is an effort to improve storage efficiency  
by detecting and removing  duplicates before data is written to the backup storage,
thus reduces extra disk space required to hold temporary data. 
This approach requires a fingerprint  comparison of the new content chunks with
previously stored fingerprints. 
As data grows to be big, content chunk lookup for fingerprint comparison from disk storage
becomes too slow to be scalable.
%and searching such disk
%However, unless some form of locality or similarity is exploited, inline, chunk-based deduplication, 
%when done at a large scale faces what has been termed the disk bottleneck problem: to facilitate fast chunk ID lookup, 
%a single index containing the chunk IDs of all the backed up chunks must be maintained. 
%As the backed up data grows, the index overflows the amount of RAM available and must be paged to disk. 
%Without locality, the index cannot be cached effectively, and it is common for nearly 
%every index access to require a random disk access. This disk bottleneck severely limits deduplication throughput.
Several techniques have been proposed to speedup searching of duplicate 
content~\cite{bottleneck08,extreme_binning09,sparseindex09}. For example,  
Zhu et al.~\cite{bottleneck08} tackle it 
by using an in-memory Bloom filter and prefetch groups of chunk IDs that are likely to be 
accessed together with high probability. It takes significant memory resource for filtering.
NG et al.~\cite{ NGmiddleware2011}  use  
a related filtering technique for integrating deduplication in Linux  file system and the memory
consumed is up to 2GB for a single machine. That is still too big in our context discussed below. 
%Lillibridge et al.~\cite{sparseindex09} break list of chunks 
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is 
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments. 
%These are segments that share many chunk IDs with the incoming segment with high probability.
%Deepavali et al.~\cite{extreme_binning09}  use signature-based file similarity  and group similar files
%into the same physical location (bins) to deduplicate against each other.

Duplicate  search approximation~\cite{extreme_binning09,sparseindex09}  has been proposed 
%in extreme bining and other sparse indexing 
to package similar content in one disk, and duplicate lookup  only searches
for chunks within files which have a similar file-level or segment-level  content fingerprints.
That leads  to a smaller amount of memory usage for storing meta data in signature
lookup with a tradeoff of the reduced recall ratio.

\section{Requirements and Design Options}
\label{sect:options}

We discuss the characteristics and 
main requirements for VM snapshot backup in a cloud environment.
which are different from a traditional data backup. 
% That arises mainly in Alibaba's Aliyun cloud service.
%concerns
%Base on Aliyun's production environment, the snapshot backup job has to satisfy 
\begin{enumerate}
\item {\em Cost consciousness.}
There are tens of thousands of VMs running on a large-scale cloud cluster. 
The amount of VM data to backup is huge and cost of complete backup can be prohibitively high without
deduplication and compression support.  Thus cost minimization is required to deliver a deplorable solution.

On the other hand, the computing resource allocated for backup is often very small.
This is because snapshot data is not retrieved frequently by users and the allocation of  cluster resource
is typically prioritized towards active services such as online web services and map-reduce jobs.
Thus  it is not easy to accept any backup behavior that affects VMs performance or stability of active users.

%snapshot service shall not compete cpu, memory, or I/O bandwidth with VMs. Specifically, the memory usage of snapshot service can never exceed 500MB in any node.
\item {\em Fast backup speed.}
The snapshot  backup  operations need to be completed quickly  every day (e.g. 1-2 hours ). 
Often a cloud service has  few hours of light workload each day (e.g. midnight),  which creates an opportunity  for backup.
But a  longer use of bandwidth and computing resource  for backup can create  noticeable  contention with the existing cloud,
which is not preferred for cloud production system operation. 
%As a large-scale cluster hosting tens of thousand of active VMs everyday,  the amount of data
%to be processed is huge. 
%For example, in an Aliyun cluster with over 1,000 nodes and each hosts over 25 VMs, The aggreated amount of data 
 % the system must finish saving daily snapshots of all VMs in 2 hours. In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 500TB/hour.

% the system must finish saving daily snapshots of all VMs in 2 hours. In our typical 1000 nodes cluster, each node hosts 25 VMs, each VM has 40GB of data on average, that translates to backup throughput of 139GB/second, or 500TB/hour.
\item {\em Fault tolerance.}
The addition of a backup service should not decrease the degree of
fault tolerance.  If data failure in  a backup service  affects many users,  such a solution is not desirable.
%when users access snapshots in a recovery process. 
\end{enumerate}

There are multiple choices in designing a backup architecture  for VM snapshots.
We discuss the following design options with a consideration on their strengths and weakness.
\begin{enumerate}
\item  {\em An external and dedicated backup storage system.} 
In this architecture setting, a separate backup storage system using
the standard backup and deduplication techniques can be deployed~\cite{bottleneck08,extreme_binning09,sparseindex09}. 
This system is attached to the cloud network and every machine can periodically transfer snapshot data to 
the attached backup system. 
A key weakness of this approach is communication bottleneck between a large number of machines
in a cloud to this centralized  service.
Another weakness is that the cost of allocating separate resource for dedicated backup  can be expensive.
Since most of backup data is not used eventually, CPU and memory resource in such a backup cluster may not be fully utilized.
\item {\em A decentralized and co-hosted backup system with full deduplication.}
In this option, the backup system runs on an existing set of cluster machines. 
The advantage is that there is no separate and dedicated cost for backup.
% and a distributed storage architecture for backup allows a possible exploitation of  data locality between
%the source of data and storage location of backup data. 
The disadvantage is that 
%the backup service would compete CPU, memory, and disk resources with the other cloud services.
even such a  backup service may only use  a fraction of the existing disk storage, 
fingerprint-based search does require a significant amount of memory for fast lookup of identical content chunks.
This competes memory resource  with the existing applications.
%Decentralized deduplication is studied in ~\cite{Clements2009} and the focus is on
%block-level copy-on-write and compare-by-value techniques.
Even approximation~\cite{extreme_binning09,sparseindex09} can be used to reduce memory requirement,  
a  key weakness of global content fingerprint comparison is that it can affect fault isolation.
Because a content chunk is compared with a content signature collected from other users,
this artificially creates data dependence among different cloud users.
% since storage for shared identical content chunks can become a failure point.
The loss of a common chunk can affect many users whose VM images share this 
chunk. Adding replicas for this global dataset can enhance the availability,
but incurs significantly more cost.
%2) data sharing among users for accessing common signagutes causes the system less resilient to storage failures.
%any loss of one piece of  shared data content hash and actualcontent will damange many VM snapshots, which can cause massive impacts
%on reliability of many users.

%Another point is that the previous work in signature-based comparison does not address
%load balancing for a distributed environment during parallel access.  
%Some content signatures can be extremely hot, but the machines  that  handle such signatures can become
% a bottleneck. Uncoordinated signature assignment could lead to imbalanced access workload.
\end{enumerate}

%There are multiple choices of snapshot backup design for VM images and our considerations are described
%as follows. 
%Our design considerations
%\begin{enumerate}
%\item {\bf Centralized vs. decenalized} 
%
%It is desirable to have  a decentralized architecture.
%Given a large amount of snapshot data communicated from each machine to the backup storage,
%with a distributed storage architecture for backup, one could exploit  exploit data locality between
%the source of data and storage location of data to reduce cross-platform bandwidth requirement for backup.

%execute in parallel and easy to coordinate. In fact, we want to avoid cross-node dependency at scheduling VM snapshot operations, such that no global coordinator is necessary.
%\item {Load balancing in resource consumption}: the cost of snapshot service shall be evenly distributed onto every node. We don't have a super powerful
%or stable node that can accept extra responsibility.
%\item {minimization of inter-user data dependency for fault tolerance}: we want to minimize the data dependency to a controllable level. Data deduplication means sharing of data, thus one failure at a single point may affect the snapshot service of hundreds of VMs, which is absolutely unacceptable.
%\item {Resource usage modeling and control}.
%\end{enumerate}

With these considerations in mind, we propose a decentralized backup architecture with multi-level and selective 
deduplication. This 
backup service is hosted   in the existing set of cloud machines and resource usage is controlled
with a minimal impact to the existing applications.
The duplication process is first conducted among data within each VM
and then is conducted  across VMs when needed to 
utilize a common dataset.  
Given the concern that searching duplicates across  VMs  is a global feature which can affect parallel performance
and complicate failure management,
we would like to use a small common dataset while still maintaining a cost-effective deduplication ratio.
For this purpose, we exploit the data characteristics of snapshots and only store most popular data.
%
%in the  our studies show that there are a large amount of data unmodified in VM snapshots
%and shared among many users (e.g. OS data). The sharing pattern follows a zip-like distribution.
%With this characteristic, we can store a small amount of popular data items which coverage a large amount of
%snapshot data blocks. 
%We discuss our design and data analysis in the next section.

%This method is based on the observation of two facts in Aliyun's VM cloud: first, VM's OS disks contain 
%huge amount of operating system and software related data which is almost never modified during VM's life span. 
%Second, the duplication pattern of user generated data follows Zipf-like law, thus a tiny set of hottest data
%take up the majority of data duplication. As a result, if we extract these hot data as a common data set,
%then most of data duplication will emerge by searching in this very small scope.

%For example, In cloud storage, we are solving data deduplication problem in a different context of 
%data stream deduplication (the D2D case): our snapshot storage service is co-located with runtime VMs,
%thus we have very limited amount of system resource leave for data deduplication. For example,
%in Aliyun's 8-core, 48GB memory, 12TB VM server, there lives over 20 VMs who are hungrily competing for
%system resources: some may running map-reduce jobs, some may serving busy web requests, 
%some may storing backend databases, any behavior that affects user VMs performance or stability is unacceptable.

%To reduce the cost of deduplication, we must confine the scope of searching duplicates as much as possible,
%thus some hints are needed to tell us where the most likely duplicates would hide. 
%Several form of hints have been used in the past: many D2D systems exploits locality,
%which bases on the fact that duplicates are likely to appear in the same sequence as they have arrived before.
%Similarity is another popular hint, it suggests that two sets of data blocks may contain lots of
%duplicates if they are indetified as similar by certain similarity detection measurements.

%\section{Architecture design and Deduplication for Snapshot Support}
%
%\section{Popularity analysis}
% data on what data popular.
%
%Analysis: modeling: singular
%
%Compute: Cache hit ratio vs. cache cost (memory cost).