### 2012-02-29 vm_snapshot_sim/2012concl.tex

\section{Conclusion}
\label{sect:final}
In this paper we propose and present a multi-level selective deduplication scheme for VM snapshots. 
with minimal cost. Inner-VM deduplication localizes backup data dependence and exposes more parallelism  
while we use  data characteristic in VM cloud backup to identify popular common data blocks
and cover  a large percentage of snapshot content blocks wit a small amount of memory index.
Our evaluation using Aliyun's implementation has shown that 
level 1 and level 2 deduplication can reduce the storage need by 80\%. 
Level 3 with small CDS memory usage can accomplish close to 44\% of what full fingerprint-based
deduplication can do and bring additional 6\% storage reduction compared to the original data size.
Noted that 6\% is still significant, which is about 24GB per each VM and for a 1000 node Aliyun cluster,
this is about 600 terabytes.

%Our experiments show th
%our solution can eliminate the majority of data duplication with a tiny fraction of
%block hash index store in memory. It does not only saves valuable system resouces in
%the VM cloud, but also makes deduplication much faster.
%
%
%Using  50 user VM data out of 1322 data disks as the training data and
%with  1.5\% as CDS threshold, we see the total 1198GB of new data is reduced by
%755.8GB, while perfect deduplication can reduce 1017.4GB. So 74.3\% of duplicate blocks are eliminated
%by pre-trained CDS, which is quite satisfactory.