### 2012-03-01 vm_snapshot_sim/main.tex
tao

\documentclass[finalversion]{usetex-v1}
%\usepackage[top=1in, bottom=1in, left=1.25in, right=1.25in]{geometry}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{epsfig}
%\usepackage{breakurl}
%% Define a new 'leo' style for the package that will use a smaller font.
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
\makeatother
%% Now actually use the newly defined style.
\urlstyle{leo}
%\linespread{1.2}
%\setlength{\parskip}{1ex}

\newcommand{\comments}[1]{}

\begin{document}
%\title{Selective Deduplication for VM Snapshots in Cloud Storage}
\title{Multi-level Selective Deduplication for VM Snapshots in Cloud Storage}
%\title{Data Deduplication and Zipf-like Distribution in VM Cloud Storage}
\author{Wei Zhang$^{\star\dagger}$, Hong Tang$^\dagger$, Hao Jiang$^\dagger$, Tao Yang$^\star$, 
Xiaogang Li$^\dagger$, and Rachael Zeng$^\dagger$ \\
   {\normalsize $^\star$UC Santa Barbara}\\
   {\normalsize$^\dagger$Alibaba.com}
}

\date{}
\maketitle

%need to explain: vm cloud, vm backup, dedup fundamentals, vm dedup(2 works), distributed dedup, zipf
%need to define: requirements, design considerations, architecture

\input{2012abs}
\input{2012intro} %introduce the vm cloud, backup problem, and our requirements
\input{2012background} %related works, related solutions, 
\input{2012arch} %, design considerations, architecture
\input{2012eval} %data study, zipf, model of cds size and dedup efficiency
%\input{results}
%\input{related}
%\input{future}
\input{2012concl}

      %\setlength{\itemsep}{ex}%
%      \setlength{\parskip}{0ex}%
%\setlength{\itemsep}{-3mm}

	%\linespread{0.3} 
  \let\oldthebibliography=\thebibliography
  \let\endoldthebibliography=\endthebibliography
  \renewenvironment{thebibliography}[1]{%
    \begin{oldthebibliography}{#1}%
      \setlength{\parskip}{-0.02ex}%
      \setlength{\itemsep}{-0.02ex}%
      \setlength{\baselineskip}{-0.02ex}%

  }%
  {%
    \end{oldthebibliography}%
  }
{\small
\bibliographystyle{abbrv}
\bibliography{dedup,dedup1}
}
\end{document}