### 2012-03-10 vm_snapshot_sim/2012abs.tex
small adjustment

\begin{abstract}
%Virtualization has became the engine behind many cloud computing platforms.
In a virtualized cloud computing environment, frequent  snapshot backup of virtual disks improves
hosting  reliability but storage demand of such operations is huge.
While dirtybit-based technique can identify unmodified data between versions, 
full deduplication with fingerprint comparison  can remove more redundant content
at the cost of computing resources.
%with  for similarity comparison and   reliability handling.
%Current snapshot deduplication is mainly done through copy-on-write 
%on fixed-size disk blocks. Such solutions cannot handle the
% cross VM data duplication because VMs do not share any data. 
%In addition, storing VM images and their snapshots
%in the same storage engine reduce the underline design flexibility because 
%these two kinds of data have distinct access requirements.
%In this paper, 
%we show that there is a large amount of duplicated data shared amongy virtual machines
%through a production VM data study and thus it is expective to perform cross-machine deduplication. 
This paper presents a multi-level selective deduplication  scheme which
integrates  inner-VM and cross-VM duplicate elimination under
a stringent resource requirement.
% minimal  resource impact to the existing cloud services.  
This scheme uses popular common data to facilitate 
fingerprint comparison while reducing the cost and it
strikes a balance  between local and global deduplication 
to  increase parallelism and  improve reliability. 
% first perform a large scale study in production VM clusters 
%to show that cross VM data duplication is severe due to they have large amount of
%common data. Then our data analysis finds out that the overall data duplication pattern follows the Zipf's law.
%Base on these discoveries, we propose a snapshot storage deduplication scheme using variable-size chunking
%to address the above problem efficiently.
%We eliminate the majority of cross VM data duplication by pre-select
%a small set of frequently seen data blocks to be shared globally, and we also remove
%many cross snapshot duplication by using smaller chunking granuarity and locality.
Experimental results  show the proposed scheme  can achieve high deduplication ratio while using
a  small  amount of cloud resources. 
\end{abstract}