### 2012-03-13 data_analysis/con.tex

\section{Conclusion}
In this paper, a empirical study using Aliyun's real user VM data traces is conducted. 
We present a number of observations and insights on the data duplication
in VM snapshot backups. We describe
in detail the patterns of VM access, VM snapshot duplication, locality reduction, and
the dintinct difference between OS and user generated data.
We hope this will facilitate future designs of cloud storage for VMs.