### 2012-03-13 data_analysis/main.tex

\documentclass[finalversion]{usetex-v1}
%\usepackage[top=1in, bottom=1in, left=1.25in, right=1.25in]{geometry}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{epsfig}
\usepackage{multirow}
\ifx\pdftexversion\undefined
\usepackage[dvips]{graphicx}
\else
\usepackage{graphicx}
\fi

\newcommand{\comments}[1]{}

\begin{document}
%\title{Selective Deduplication for VM Snapshots in Cloud Storage}
\title{Understanding the Data Duplication in VM Cloud Storage}
%\title{Data Deduplication and Zipf-like Distribution in VM Cloud Storage}
\author{
Wei Zhang$^{\star\dagger}$, Hong Tang$^\dagger$, Hao Jiang$^\dagger$, Tao Yang$^\star$\\
   {\normalsize $^\star$UC Santa Barbara, \{wei, tyang\}@cs.ucsb.edu}\\
   {\normalsize$^\dagger$Alibaba.com, \{hongtang, haojiang\}@alibaba.com}
}
\date{}
\maketitle

\input{abs}
\input{intro} %introduce the vm cloud, snapshot
\input{overall} %complete dedup on both datasets 
\input{locality} %
\input{user}
\input{dup}
%\input{count} %data study, zipf, model of cds size and dedup efficiency
%\input{results}
%\input{related}
%\input{conclusion}
\input{con}

      %\setlength{\itemsep}{ex}%
%      \setlength{\parskip}{0ex}%
%\setlength{\itemsep}{-3mm}

	%\linespread{0.3} 
%  \let\oldthebibliography=\thebibliography
%  \let\endoldthebibliography=\endthebibliography
%  \renewenvironment{thebibliography}[1]{%
%    \begin{oldthebibliography}{#1}%
%      \setlength{\parskip}{-0.02ex}%
%      \setlength{\itemsep}{-0.02ex}%
%      \setlength{\baselineskip}{-0.02ex}%

%  }%
%  {%
%    \end{oldthebibliography}%
%  }
{\small
\bibliographystyle{abbrv}
\bibliography{dedup1,library}
}
\end{document}