### 2012-03-18 data_analysis/con.tex
update

\section{Conclusion}
In this paper, a empirical study using Aliyun's real user VM data traces is conducted. 
We present a number of observations and insights on where and how data duplication
exist in VM snapshot backups. We describe
in detail the patterns of VM access, the potential of deduplication, 
important factors such as locality, similarity and popularity, and
the dintinct difference between OS and user generated data duplication.
We hope this study will facilitate researchers and engineers to design better
data deduplication strategies
for backup storage in VM Cloud.