### 2012-08-25 snapshot_system/main.tex

\documentclass{acm_proc_article-sp}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{epsfig}
\usepackage{breqn}

\begin{document}

\title{An In-Cloud Backup System With Deduplication for VM Snapshot}

% \numberofauthors{7}
% \author{
% \alignauthor Wei Zhang\\
%        \affaddr{CS Dept.}\\
%        \affaddr{UC Santa Barbara}\\
%        \email{wei@cs.ucsb.edu}
% \alignauthor Hong Tang\\
%        \affaddr{Aliyun Inc.}\\
%        \affaddr{Alibaba Group}\\
%        \email{hongtang@alibaba-inc.com}
% \alignauthor Tao Yang\\
%        \affaddr{CS Dept.}\\
%        \affaddr{UC Santa Barbara}\\
%        \email{tyang@cs.ucsb.edu}
% }

\maketitle

\input{abs}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

\input{intro}
\input{design}
\input{snapshot}
\input{append}
\input{cds}
\input{failover}
\input{result}
\input{con}

\bibliographystyle{abbrv}
\bibliography{library}
\end{document}