### 2012-12-20 snapshot_system/challenge.tex
paper update

\section{Challenges}
\subsection{Indexing}
Most deduplication systems operate at the sub-file level:
a file or a data stream is divided into a sequence of fixed
or variable sized segments. For each segment, a cryptographic
hash (MD5, SHA-1/2, etc.) is calculated as its
fingerprint (FP), and it is used to uniquely identify that
particular segment. A fingerprint index is used as a catalog
of FPs stored in the system, allowing the detection
of duplicates: during backup, if a tuple of the form <
FP, location > exists in the index for a particular
FP, then a reference to the existing copy of the chunk
is created. Otherwise, the chunk is considered new, a
copy is stored on the server and the index is updated accordingly.
In many systems, the FP index is also crucial
for the restore process, as index entries are used to locate
the exact storage location of the chunks the backup
consists of.

The index needs to have three important properties:
1) scale to high capacities, 2) achieve good indexing
throughput, and 3) provide high duplicate detection
rate—i.e., high deduplication efficiency. Table 1 demonstrates
how these goals become very challenging for a
Petascale system. Consider a typical virtualized cloud 
in which 10,000 VMs store 1 PB data, and the
chunk size is 4 KB (for fine-granularity duplicate detection),
indexing capacity will need to be at least 10,000
GB to support all 250 billion objects in the system. Such
an index is impossible to maintain in memory. Storing it
on disk, however, would greatly reduce query throughput.
To support 1000 concurrent snapshot backup tasks, 
an aggregate throughput of 20 GB/sec, would require the
index and the whole dedupe system for that matter
to provide a query service throughput of at least 5,000
Kops/sec. Trying to scale to 1 PB by storing the index
on disk would make it impossible to achieve this level
of performance. Making the chunk size larger (e.g.,
128 KB) would make deduplication far more coarse and
severely reduce its efficiency.

Even we can distribute the index over the cloud machines,
the resources that can be allocated to system services are
limited to minimize the negative impact to user experiences. 
So it becomes obvious that efficient, scalable indexing is
a hard problem. On top of all other indexing challenges,
one must point out that chunk FPs are cryptographic
hashes, randomly distributed in the index. Adjacent index
entries share no locality and any kind of simple readahead
scheme could not amortize the cost of storing index
entries on disk.

\subsection{Deletion}
Contrary to a traditional backup system, a dedupe system
shares data among files by default. 
Since snapshots in VM cloud face constant deletion, 
reference management
is necessary to keep track of chunk usage and reclaim
freed space. In addition to scalability and speed, reliability
is another challenge for reference management.
If a chunk gets freed while it is still referenced by snapshots,
data loss occurs and files cannot be restored. On the other
hand, if a chunk is referenced when it is actually no
longer in use, it causes storage leakage.

For simplicity, many previous works only investigated simple reference
counting without considering reliability
and recoverability. Reference counting, however, suffers
from low reliability, since it is vulnerable to lost or repeated
updates: when errors occur some chunks may
be updated and some may not. Complicated transaction
rollback logic is required to make reference counts consistent.
Moreover, if a chunk becomes corrupted, it is
important to know which files are using it so as to recover
the lost segment by backing up the file again. Unfortunately,
reference counting cannot provide such information.
Finally, there is almost no way to verify if the
reference count is correct or not in a large dynamic system.
Our field feedback indicates that power outages and
data corruption are really not that rare. In real deployments,
where data integrity and recoverability directly
affect product reputation, simple reference counting is
unsatisfactory.

Maintaining a reference list is a better solution: it is
immune to repeated updates and it can identify the files
that use a particular segment. However, some kind of logging
is still necessary to ensure correctness in the case of
lost operations. More importantly, variable length reference
lists need to be stored on disk for each chunk.
Every time a reference list is updated, the whole list (and
possibly its adjacent reference lists—due to the lists’
variable length) must be rewritten. This greatly hurts the
speed of reference management.

Mark-and-sweep is generally a better solution. 
During the mark phase, all snapshot metadata are traversed 
so as to mark the used chunks. 
In the sweep phase all chunks are swept and unmarked chunks are reclaimed. 
This approach is very resilient to errors: 
at any time the process can simply be restarted with no negative side effects. 
Scalability, however, is an issue.
To deal with 1 PB snapshots data, which may be represented by 10,000 GB FP index,
If we account
for an average deduplication factor of 10 (i.e., each chunk
is referenced by 10 different snapshots),
the total size
of snapshot metadata that need to be read during the mark phase will
be 100 TB. 
This alone will take almost 2 hours on a
1000-machine cloud. Furthermore, marking the in-use
bits for 250 billion entries is no easy task. There is no
way to put the bit map in memory. One might want
to mitigate the poor performance of mark-and-sweep by
doing it less frequently. But in practice this is not a viable
option: customers always want to keep the utilization
of the system close to its capacity so that a longer
history can be stored. With daily snapshots taking place,
systems rarely have the luxury to postpone deletion operations
for a long time. In our field deployment, deletion
is done once a day. More than 2 hours in each run is too
much. In a large production-oriented dedupe system reference
management needs to be very reliable and have
good recoverability. It should tolerate errors and always
ensure correctness. Although mark-and-sweep provides
these properties, its performance is proportional to the
capacity of the system, thus limiting its scalability.

\subsection{Fault Tolerance}
Because deduplication storage system
 artificially creates data dependency among different VM users,
it increases the risk of data availability.
In large scale cloud, node failures happen at daily basis,
the loss of access to a shared chunk can affect many VMs whose snapshots share this chunk. 
Without any control of the scope of data sharing, we can only increase the level of 
replication to enhance availability, which would significantly 
balance out the benefits of deduplication.