### 2013-03-10 hotstorage13/arch.tex
new

\section{System Design}
\label{sect:arch}

%\begin{figure}[htbp]
%  \centering
  %\epsfig{file=images/arch.eps, height=2in, width=2.66in}
%  \epsfig{file=images/arch.eps, width=3.9in}
%  \caption{Snapshot backup architecture of each node.}
%  \label{fig:arch}
%\end{figure}

We consider deduplication in two levels. The first level
uses coarse-grain segment  dirty bits for version-based detection~\cite{Clements2009,Vrable2009} to 
identify difference from the previous snapshot to the current snapshot.  
%That  can be accomplished efficiently and inexpensively in the OS level.
%We use a segment-level coarse-grain setting to reduce the efforts in detection and maintaining dirty bits. 
Our experiment with Alibaba's production dataset shows that over 70 percentage of 
duplicates can be detected using segment dirty bits when the segment size is 2M bytes.  
This setting requires OS to keep the metadata for segment dirty bits and the amount of space for such segment 
dirty bits is negligible. In the second level of deduplication, content blocks of dirty segments at each VM 
are compared with the fingerprints of unique  blocks from the previous snapshots.
Our key strategies are explained as follows.
\begin{itemize}

\item {\bf Separation of duplicate detection and data backup.}
%Request accumulation and partition-based deduplication.}
The second level duplicate detection requires a global comparison of 
content fingerprints.
% of stored chunks with the new chunks scanned during the backup process. 
%Bloom filters and caching allows some of content 
%fingerprints to be stored on the cheap disks~\cite{bottleneck08}. Such an optimization is good 
%for inline deduplication where chunks scanned need to be determined to be duplicate or not instantly without 
%waiting while  it memory consumption is still significant in competing with stand virtual 
%machine activities.  
Our approach is to perform duplicate detection first before actual data backup.
That requires a pre-scanning of  data  blocks in dirty VM segments  and after that, real backup  for those non-duplicates
is conducted. That does incur an extra  round of local VM block reading while avoiding inline 
deduplication and leading to a much smaller resource requirement.  
During VM reading, detection requests can be accumulated on the disk 
for a period time  in a  partitioned manner.  Aggregated duplicate requests can be processed partition by partition. 
%We also accumulate the delete requests and perform them partition by partitions. 
Since each partition corresponds to a small portion of global content index, the cost to 
process detection requests within a partition is significantly smaller than processing all the requests globally.

\item {\bf Buffered data redistribution in parallel duplicate detection}.  
Let {\em global index} be the meta data containing the fingerprint values of unique blocks of snapshots
in all VMs and  the reference pointers which lead to the location of raw data.
%A duplicate detection request for a content block needs to be compared with the global index to determine if this
%block is non-duplicate. If it is a duplicate, return the corresponding reference pointer.
%We conduct parallel processing of detection requests in all machines involved. 
A logical way to distribute detection requests among machines is to partition based on fingerprint values of content blocks
in a  disjointed manner.
Initial VM data reading follows the VM distribution among machines and the detected duplicate summary results 
need to be collected following the VM distribution. 
Therefore, there are two all-to-all data redistribution operations involved.
One is to map detection requests from VM-based distribution to fingerprint based distribution.  
One is to map duplicate summary from fingerprint-based distribution to VM based distribution.  
The redistributed data needs to be accumulated on the disk to minimize the use of memory.
To minimize the excessive disk seek cost, outgoing or incoming data exchange messages need to be buffered to bundle small messages.
Given there are $p\times q$ partitions where $p$ is the number of machines and $q$ is the number of fingerprint-based partitions
at each machine, space per each buffer  is small under the memory constraint for large $p$ or $q$ values.
This counteracts the effort of seek cost reduction.  
%Then there will be a large number of storage  IO operations involved, suffering the huge seek overhead.
We have designed an efficient data exchange and disk data buffering  scheme to address this.

%As we detect and backup requests before actual backup takes place,  we need to store chunk data on 
%temporary disks and once duplicates 
%are detected, we only need to fetch and store these non-duplicates 
%in the backup storage.  To minimize the storage usage of temporary disk space, we donot save the data content 
%of accumulated requests with replication, but we store  data separately for each virtual machine. In an event 
%that the storage of temporary accumulation fails for certain virtual machines, rescanning of these virtual machine images 
%is conducted and backup of these machines is re-initiated.   Meta data such as content hash for accumulated requests 
%is stored separately since we map the meta data into a set of buckets using chunk fingerprints. 
%The benefit of separating content and meta data is to allow fast recovery of backup operations while enabling bucket-based lazy deduplication.

\end{itemize}

We assume a flat architecture in which  all $p$ machines that host VMs in a cluster can 
be used in parallel for deduplication. 
A small amount of local disk space and memory on each machine can be used 
to store global index and temporary data. 
The real backup storage can be either a distributed file system built on
this cluster  or use another  external storage system. 
%Our design is to minimize the usage of local memory and storage on each machine.

%While we assume all machines can process  independent or coordinated
%deduplication and backup operations in parallel,  
%our scheme uses a single-thread procedure on each machine to minimize the use of CPU, network bandwidth, and disk bandwidth.

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{snapshotdata.pdf}
\caption{ Metadata structure of a VM snapshot.}
\label{fig:snapshot}
\end{figure}
%\subsection{Snapshot Representation}

The representation of each snapshot in the backup storage
has a two-level index structure in the form of a hierarchical
directed acyclic graph as shown in Figure~\ref{fig:snapshot}.
An VM image is divided into a set of segments and each  segment contains hundreds of content blocks. 
These blocks are of variable-size, partitioned using
the standard chunking technique~\cite{similar94} with 4KB as the average block size. 
%To simplify the deduplication process, segments are aligned to fix-sized boundaries, currently using 2MB.
The snapshot metadata  contains a list of segments and other meta data information.
Segment metadata  contains its  content block fingerprints and reference pointers. 
If a segment is not changed from one snapshot to another, indicated by a dirty bit embedded in the virtual disk driver, 
%its content blocks are not changed as well, thus 
its segment metadata contains a reference pointer to an earlier segment.
For a dirty segment, if one of its blocks is duplicate to another block in the system,  
the block metadata contains a reference pointer to the earlier block.

\begin{figure*}[tbhp]
\centering
\includegraphics[width=0.75\textwidth]{steps.pdf}
\caption{Processing flow of Stages  1,  2 and 3.}
\label{fig:flow}
\end{figure*}

%\subsection{Processing flow and resource usage}
The data flow of our multi-stage duplicate detection is depicted in Figure~\ref{fig:flow}. 
%let $v$ be the total number of virtual machines hosted in  a cluster,  
%we divide the backup into $k$  iterations. 
%\begin{itemize}
%\item
In Stage 1, each machine independently reads  
%$v/k$  
VM images that need a backup
and forms duplicate  detection requests. For those segments that have been modified since last backup indicated by 
dirty bits,  the system divides  each segment into a sequence of chunk blocks,  computes the meta 
information such as chunk fingerprints,  sends a request to a proper machine, accumulates  
received requests into a partition on the local temporary disk storage. 
The partition mapping is using a hash function applied to the content fingerprint. 
Assuming all machines have a  homogeneous resource configuration, each machine is evenly  assigned with
$q$ parititons of content hash and accumulates corresponding requests on the disk. 
There are two options to assign buffers in each machine to buffer requests for each partition.
1) Each machine has  $p\times q$ send buffers corresponding to $p\times q$ partitions in the cluster
since a content block in a VM image of this machine can be sent to any of these partitions.
2) Each machine allocates $p$ send buffers to deliver requests to $p$ machines and 
$p$ receive buffers to collect requests  from other machines.
Then the system copies requests from each of $p$ receive buffers to  $q$ local request buffers,
and outputs each request buffer to one of the request partitions on the disk
when this request buffer becomes full.  Option 2 is much more efficient than Option 1 because $2p+q$ is much smaller than
$p\times q$. As a result, each buffer has a bigger size to accumulate requests and that means
less disk seek overhead.

%We assume that the startup cost for network message sending is  much more less (e.g. an order of magnitude
%less) than from one machine to another 
%disk storage startup cost such as seek is much mu
%Note that we only accumulate requests with their meta data information.
%\item
Stage  2 is to load disk data and perform fingerprint comparison at each machine one request partition at a time.
% in the partitions.  
%The system maintains the index of all chunk fingerprints divided in the buckets. 
%The system loads the global index partition 
%and accumulated corresponding requests, and compare them to  identify the duplicated blocks.  
%Then we unload them, load the global index and accumulated requests for another partition. 
%This process is repeated until all buckets are processed.
At each iteration, once in-memory comparison between an index partition and request partition is completed,  
duplicate summary information for segments of each VM is routed from the fingerprint-based distribution  to the
VM-based distribution.  The summary contains the block ID and  the reference pointer for each detected duplicate block.   
Each machine uses memory space of the request partition as a send buffer with no extra memory requirement.
But it needs to allocate $p$ receive buffers to collect duplicate summary from other machines.
It also allocates $v$ request buffers to copy duplicate summary from $p$ receive buffers and output to the local disk
when request buffers are full.

%. In this way, it sends duplicate summary to the corresponding machines one by one to minimize the startup cost for sending a message.
%In the receiving side, 
%All-to-all machines data exchange is conducted  so that  
%each machine allocates $q$ receive buffers corresponding to
%$q$ local partitions, to hold received messages from other machines.  
%When a receive buffer for a VM is  full, the data is written to the local disk storage. 
%These substeps are repeated as the stream of data is exchanged when each machine compares through all $q$ partitions.

%\item 
Stage 3 is to perform real backup one VM at a time.
The system loads the duplicate summary of a VM, 
reads  dirty segments of a VM, and output non-duplicate blocks to the final backup 
storage. Additionally, the global index on each machine is updated with the meta data of new chunk blocks. 
When a segment is not dirty, the system only needs to output the segment meta data such as a reference pointer. 
There is an option to directly read dirty blocks instead of fetching a dirty segment which can include duplicate
blocks. Our experiment shows that it is faster to read dirty segments in the tested workload.
Another issue is that during global index update, the system  may find some  blocks with the same fingerprint may be 
created redundantly. For example, two different VM blocks have the same content block fingerprint while the  global 
index does not have such a fingerprint. The redundancy is discovered and logged during the index update and can be repaired
periodically when necessary.  Our experience is that there is a redundancy during the initial snapshot backup and once 
that is repaired, the percentage of redundant blocks due to concurrent processing  is insignificant.

%If a segment is dirty,  the system  use the duplicate summary information from Step 2 for internal 
%duplicate blocks of this dirty segment and output the remaining non-duplicate blocks to the backup storage.
%\end{itemize}

%          In fetching non-duplicates,    it is sometime faster that we fetch a large number of consecutive chunk blocks from an accumulated  segment which contains duplicated chunks,  to avoid the random seek overhead.

The above steps can be executed by each machine
using one thread to minimize the use of CPU, network bandwidth, and disk bandwidth.
The  disk storage usage on each machine 
is fairly small for  storing part of global index and
accumulating  duplicate detection requests that contain fingerprint information.   
We impose a memory limit $M$ allocated for each stage of processing at each machine.
The usage of $M$ is controlled as follows and the distribution among buffers is optimized based on the relative
ratio between the cross-machine network  startup cost and disk access startup cost such as seek time.
Slower start cost requires more buffer space.
\begin{itemize}
\item For Stage 1, $M$  is divided for 
1) an I/O buffer to read dirty segments; 2) $2p$ send/receive buffers and $q$ request  buffers.

%1) $p$ send buffers. images, sends the fingerprint of a dirty block to a machine that is responsible for this fingerprint.
%2) Buffer requests gathered from other machines and divide those requests into $q$ buckets. Once a buffer for each bucket is full, output to the disk storage.
\item 
For Stage 2,  $M$  is divided for 1) space for hosting a global index partition and 
the corresponding request partition; 2) $p$ receive buffers and $v$ summary buffers.
%%accumulating duplicate summary for each targeted VM.   Once  each machine processes a bucket of  duplicate detection requests, 
%For each machine which hosts $v$  VMs, and will receive messages from many other machines, 
%the system needs to allocate buffer space for each VM ot accumulate the duplicate summary records. 
%When a buffer for a VM is full, the result is written to the disk.

\item For Stage 3, $M$  is divided for 1) an I/O buffer to read dirty segments of a VM and   
write non-duplicate content blocks to the  backup storage.
2) summary of duplicate blocks within dirty segments of  a VM.
\end{itemize}
%We find the memory for step 2 is most critical for buffering duplicate summary in hiding the latency. 
%Certainly most storage system can allow many IO requests issued in parallel overlapping the IO latency.  
%Here we assume the worst case and also minimize the IO bandwidth usages by issuing 
%one request at a time to use the minimal disk bandwidth resource.

{\bf Snapshot deletion.} Each VM will keep a limited number of  automatically-saved VM snapshots and thus
expired snapshots will be deleted unless its owner explicitly asks for its retention.
We adopt the idea of mark-and-sweep~\cite{Guo2011}. 
A block or a segment can be deleted if its reference count is zero.
%As we discussed in Section~\ref{sect:data}, the reference counter is kept in the global index, partitioned among machines. 
To delete useless  blocks or segments periodically, we read the meta data  of live 
snapshots and compute the reference count of all blocks and segments  in parallel
while following the fingerprint-based partitioning.
Similar to the multi-stage duplicate detection process, reference counting is conducted in multi-stages. 
Stage  1 is to read  the segment and block  metadata of snapshots 
to accumulate  reference count requests in different machines  in the fingerprint based distribution.
Stage   2 is to count references within each partition and detect those records with zero 
reference. 
%Stage 3 is that every VM gathers its z confirmed deletion requests for its content and sends them to the backup storage.
The backup  data repository logs deletion instructions,  and will periodically perform a compaction operation when 
its deletion log is too big. 
%Due to the paper constraint, we will not discuss the deletion operations in details.

%The above method requires the counter information to be consistently maintained in the global index.
%Finally the deleteion summary. it will adjust the corresponding reference counts. At the end of bucket worker's operation, it outputs the unused blocks information to each VM data repository's deletion log.

%This compaction will make a copy of all the data in use, and remove the obsoleted blocks. However, after such a compaction, many block offsets will be changed, as a result, such changes are sent to the bucket job queue so that changes will be reflected upon the next bucket update.

%Since we maintain a single copy of each block information in the bucket index, and there is at most one bucket worker operating the bucket index at one time, it it safe to do the simple referencec ounting rather than mark-and-sweep.