### 2013-03-10 hotstorage13/background.tex

\section{Background and Related Work}
\label{sect:background}

%In a virtualized cloud environment such as ones provided by Amazon EC2\cite{AmazonEC2} and Alibaba Aliyun\cite{Aliyun}, 
At a cloud cluster node, each instance of a guest operating system runs on a virtual machine, accessing virtual hard disks 
represented as virtual disk image files in the host operating system.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations are taken place at the virtual device driver level, which means no fine-grained file system metadata can be used to determine the changed data. Only raw access information at disk block level are provided. 
%Each physical machine hosts many VMs and petabytes of data in a cloud cluster need a frequent  backup. 
%Ideally speaking, snapshot backup must not affect the normal cloud service, which means that 
%only a very small slice of cluster resource can be used for the backup purpose.

%The previous work for storage backup has extensively used  data deduplication techniques can eliminate redundancy globally among different files from different users.
Backup systems have been developed to use content hash (fingerprints) to identify duplicate
content~\cite{venti02,Rhea2008}.
Offline deduplication is used in ~\cite{EMC,NetAppOffline} to remove previously written duplicate blocks during idle time.
%,NGmiddleware2011}.
%Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper}
%use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
Several techniques have been proposed to speedup searching of duplicate
content. For example,
The data domain method ~\cite{bottleneck08} 
uses  in-memory Bloom filter and maintains a prefetch cache for data blocks  which may be
accessed with high probability.  An improvement to this work with parallelization is in ~\cite{DEBAR}.
Still such methods use  a significant amount of memory resource for filtering and caching.
%NG et al.~\cite{ NGmiddleware2011}  use
%a related filtering technique for integrating deduplication in Linux  file system and the memory
%consumed is up to 2GB for a single machine. That is still too big in our context discussed below.
%Lillibridge et al.~\cite{sparseindex09} break list of chunks
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments.
%These are segments that share many chunk IDs with the incoming segment with high probability.
The approximation techniques are studied in~\cite{extreme_binning09,WeiZhangIEEE}  
%Deepavali et al.~\cite{extreme_binning09} and Zhang et al.~\cite{WeiZhangIEEE}  
to reduce memory requirement with a tradeoff of the reduced deduplication ratio.
%use fingerprint-based file similarity  and group similar files into the same physical location (bins) to deduplicate against each other.
%That leads  to a smaller amount of memory usage for storing meta data in fingerprint
%lookup  
In comparison, this paper focuses on  full deduplication without approximation.

Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,idedup}. 
All of the above approaches have focused on
such inline duplicate detection in which  deduplication of an individual block  is on the critical write path.
In our work, this constraint is relaxed and 
there is a waiting time for many duplicate detection requests. This relaxation is acceptable because 
in our context, finishing the backup of required VM images within a reasonable time window is more
important than optimizing individual VM block  backup requests.

%It should be emphasized that our approach does not  accumulate raw backup data temporarily for deduplication
%and does not require a significant amount of extra storage capacity.
%Our strategy is to perform a drity  scan to collect
% a small amount of disk space  to accumulate
%deduplicate requests along with necessary meta data, and perform actual backup after deduplicate detection completes.

%affecting the overall time of backup.