### 2013-05-03 hotstorage13/main.tex

\documentclass[finalversion]{usetex-v1}
%\usepackage[top=1in, bottom=1in, left=1.25in, right=1.25in]{geometry}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{epsfig}
%\usepackage{breakurl}
%% Define a new 'leo' style for the package that will use a smaller font.
%\makeatletter
%\def\url@leostyle{%
%  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
%\makeatother
%% Now actually use the newly defined style.
%\urlstyle{leo}
%\linespread{1.2}
%\setlength{\parskip}{1ex}

\newcommand{\comments}[1]{}

\begin{document}
%\title{Selective Deduplication for VM Snapshots in Cloud Storage}
\title{ Low-Cost Data Deduplication  for  Virtual Machine Backup in Cloud Storage}
%\title{Data Deduplication and Zipf-like Distribution in VM Cloud Storage}

\author{
Wei Zhang$^{\star}$, Tao Yang$^\star$, Gautham Narayanasamy$^\star$, Hong Tang$^\dagger$ \\
{\normalsize$^\star$ University of California at Santa Barbara}, {\normalsize$^\dagger$Alibaba Inc.}
}

% \{wei, tyang\}@cs.ucsb.edu

\date{}
\maketitle

%need to explain: vm cloud, vm backup, dedup fundamentals, vm dedup(2 works), distributed dedup, zipf
%need to define: requirements, design considerations, architecture

\input{abs}
\input{intro} %introduce the vm cloud, backup problem, and our requirements
\input{background} %related works, related solutions, 
\input{arch} %, design considerations, architecture
%\input{analysis} %, design considerations, architecture
\input{eval} %data study, zipf, model of cds size and dedup efficiency
%\input{results}
%\input{related}
%\input{future}
\input{concl}

{\bf Acknowledgment.} We thank the anonymous referees for their valuable comments.
This work is supported in part by NSF IIS-1118106. 
Any opinions, findings, conclusions or recommendations expressed in this material
are those of the authors and
do not necessarily reflect the views of the National Science Foundation.

      %\setlength{\itemsep}{ex}%
%      \setlength{\parskip}{0ex}%
%\setlength{\itemsep}{-3mm}

	%\linespread{0.3} 
%  \let\oldthebibliography=\thebibliography
%  \let\endoldthebibliography=\endthebibliography
%  \renewenvironment{thebibliography}[1]{%
%    \begin{oldthebibliography}{#1}%
%      \setlength{\parskip}{-0.02ex}%
%      \setlength{\itemsep}{-0.02ex}%
%      \setlength{\baselineskip}{-0.02ex}%

%  }%
%  {%
%    \end{oldthebibliography}%
%  }
{\small
\bibliographystyle{abbrv}
\bibliography{library,dedup1}
}
\end{document}