### 2013-05-24 snapshot_system/cds.tex
Add cds analysiswq

\section{CDS Analysis}
[Purpose of using CDS]
CDS is a small percentage of the global data index that we into DHT as a shared index. 
With CDS as as a public data resource, each VM's snapshot backup has no need to 
store its own copies for data that already exist in CDS, instead they can just
store a reference. 
We expect CDS to contain most of the data related to operating system, software and libraries.

[What's the benifits]
Using CDS as a common data source brings us several benefits: 
First, the data we collect into CDS are commonly used in many VMs, 
thus the CDS index, a special subset of global index, has very high deduplication efficency.
Second, in our architecture the CDS data are the only data that can be shared between different VMs,
by carefully control the replication degree of CDS data, we could improve the fault-tolerancy
in such a data sharing environment. Finally, the CDS index is small enough to be put into a 
in-memory DHT, thus it avoids the excess memory usage of global index 
while brings us the majority of deduplication efficiency of it.

[Describe how CDS is generated]

[Explain why CDS scales effectively as data grow]
Our empirical study based on VM images from production environment\cite{ieeecloud} showed that the 
frequency of data duplication follows Zipf-like distribution\cite{zipf}, 
with the exponent $\alpha$ between 0.65 ~ 0.70. 
As a result, it can be proved that deduplication efficiency of CDS index is scalable:

[Describe the dedup efficiency model in detail]
Let $S_c$ be the total number of FPs in CDS, $S_g$ be the total number of FPs in global index, $F$ be the size of a FP, $N$ be the number of nodes in the cluster, $m$ be the memory on each node that are used by CDS, $D$ be the amount of data on each node, and $B$ be the average block size. The deduplication efficiency will be:
\begin{equation}
  E = (S_c / S_g)^{1-\alpha}
\end{equation}
The $S_c$ and $S_g$ can be expressed as:
\begin{equation}
S_c = N*m/F, S_g = N*D/B
\end{equation}
By replacing $S_c$ and $S_g$ in the first formula, we have:
\begin{equation}
  E = (\frac{m*B}{F*D})^{1-\alpha}
\end{equation}

Since $B$, $D$ and $F$ are pre-configured constants, the deduplication efficiency of CDS is only controlled by the its memory usage.