### 2013-05-28 snapshot_system/main.tex

\documentclass{acm_proc_article-sp}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{epsfig}
\usepackage{breqn}

\newcommand{\comments}[1]{}

\begin{document}

%\title{Deduplication Localization for Virtual Machine Snapshot Backup}
\title{Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup}
%\title{BigArchive: Low-cost and Scalable Deduplication Storage for VM Snapshot}

\author{
Wei Zhang, Michael Agun,  Tao Yang \\
%Wei Zhang$^{\star}$, Michael Agun,  Tao Yang$^\star$
%{\normalsize$^\star$  University of California at Santa Barbara}, {\normalsize$^\dagger$Alibaba Inc.}
%Wei Zhang, Tao Yang, Gautham Narayanasamy\\
Department of Computer Science\\
University of California,  Santa Barbara, CA 93106
%\and
%Hong Tang\\
%Alibaba Inc.
}
%, Gautham Narayanasamy$^\star$, and Hong Tang$^\dagger$ \\

\maketitle

\input{abs}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Theory}

%\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

\input{intro}
%\input{design}
\input{background}
\input{challenge}
\input{snapshot}
\input{failover}
\input{design1}
%\input{snapshot}
%\input{append}\
\input{cds}
%\input{delete}
%\input{failover}
\input{result}
\input{con}

\bibliographystyle{abbrv}
\bibliography{library}
\end{document}