### 2013-06-20 snapshot_system/result.tex
Iadd experiment results

\section{Experiments}
Our main test-bed is an cluster of 6 machines,
each of which is equiped with a 8-core AMD FX-8120 at 3.1 GHz
with 16 GB RAM, running Linux. 
A distributed file system (QFS) runs

We used two main data sets for testing. Our synthetic
data set consists of multiple 3 GB files, each with globally
unique data segments. Our second data set consists
of virtual machine images, which are a very common
real-world enterprise use-case, that takes advantage
of deduplication. We use a VMware “gold” disk image
(VM0), hosting a Microsoft Windows XP installation,
and created three additional versions of it (VM1, VM2,
and VM3), each with incremental changes: VM1 is VM0
with all Microsoft updates and service packs, VM2 is