### 2013-06-24 snapshot_system/result.tex
Add performance analysis

\section{Experiments}
Our main test-bed is an cluster of 6 machines,
each of which is equiped with a 8-core AMD FX-8120 at 3.1 GHz
with 16 GB RAM, running Linux.
A distributed file system (QFS) runs in the cluster manages six 3TB disks
 with default replication degree set to 3.
Our data set consists of 350 virtual machine image snapshots taken from Alibaba.com's
public VM cloud in China. We select 35 VMs from the most popular 7 OSes: 
Debian, Ubuntu, Redhat, CentOS, Win2003 32bit, win2003 64 bit and win2008 64 bit. 
For each OS, 5 VMs are chosen, and every VM come with 10 full snapshots.

\subsection{System Performance}
{\bf Single VM Backup} We start the evauluation of our system by examine the normal warmed-up
backup scenario - taking a snapshot of a VM which already has old backups exist. To simulate this
scenario, we pick one 40GB VM from the data set, make an initial snapshot backup, then
evaulate the system performance on the second snapshot backup. We repeat such procedure
under different CDS memory usage settings, the results of backup time break down are shown in
firure~\ref{fig:single_vm_backup}.

It's easy to see that the time of I/O dominates the backup process, our deduplication
procedure only takes a tiny fraction of the total backup time. Larger memory permits us to
use larger CDS index to cover more of global index, as a result the amount of data to be writen
is reduced and so do the writing time.
Inside the deduplication procedure, the time costs of level-1
is plainly zero, and the costs of level-2 and level-3 are almost identical under different settings
because the number of fingerprints to process are just the same.
\begin{figure}[htbp]
  \centering
  \epsfig{file=figures/single_backup_time.pdf, width=3.5in}
  \caption{Time to backup a 40GB VM snapshot under different CDS memory usage settings}
  \label{fig:single_vm_backup}
\end{figure}

{\bf Parallel VM Backup} We then evaluate the performance of writing snapshots in parallel.
To begin, on each node we let 4 VMs writing snapshot concurrently, and gradually 
increase number of VMs to 12 to saturate our system capability. We observed 
the per-node throughput peaked at 2700 MB/s when writing 10 VM snapshots in parallel, 
which is far beyond our QFS file system capability. The reason behind it is our efficient
deduplication architecture and compression greatly reduce the amount of data that need to be writen to
the file system. The main bottleneck here is our QFS only manages one disk per node, 
making it inefficient to utilize the benefits of parallel disk access. We expect our achitecture can
perform even better in production cluster which normally has more than ten disks on each node.

\begin{figure}[htbp]
  \centering
  \epsfig{file=figures/parallel_backup.pdf, width=3.5in}
  \caption{Throughput per-node with concurrent snapshot backup tasks}
  \label{fig:parallel_backup}
\end{figure}

\subsection{Read Snapshot}

\subsection{Comparison with Other Approaches}
\subsubsection{Sampled Index}
One alternative approach to avoiding a single global index is to use a sampled
index. This is the approach taken by Guo and Efstathopoulos\cite{Guo2011}. We
compare their solution to ours by running their algorithm on our 350 traces
using a sampling rate of 1 out of 101 blocks, and always including the first
block in a container (which we fix at 16MB in size). The results of running this
test are shown in Table~\ref{??}, and the memory usage comparison can be found
Sec.~\ref{??}. Our results show that using a sampled index achieves a very high
rate of deduplication, and so a sampled index might be a good way to do dedup
in a single node setup.

The problems for sampling are in distributing the
algorithm and in deduplicating extremely large bodies of data. The algorithm
stores the entire (sampled) index in memory, which is required for high
throughput to avoid the disk-bottleneck problem - since the index itself has no
locality information, lookups are essentially random, so the index must be
somehow optimized to prevent excessive disk seeking (the sampling algorithm does
this by storing the index in memory). To distribute this index, each node would
either require a complete copy of the index which would have a prohibitive cost in
memory, or something like memcached must be used, which leaves the same problem
as the disk bottleneck with every index check potentially using the network.
The difficulty in storing extremely large bodies of data
is related, as once 1PB is stored in the system, assuming a sampling rate of 
1/101 with 22 byte index entires, 55GB of (in memory) index are required.

Our solution uses more total memory (??how much??), but is more scalable both in
terms of capactity and distributing the deduplication. The only part of our
algorithm which requires significant network traffic is the CDS deduplication,
but this is done after Levels 1 and 2 (dirty bits and comparison to parent), and
so is only done for a small percent of blocks (?? 5-10\% ??).

\subsubsection{Scalable Data Routing}
Another approach to avoiding a global index is to use a content-based hash
partitioning algorithm to break up the deduplication work. This approach is
taken by Dong et al. in their Scalable Data Routing paper, and is similar to
Exreme Binning\cite{??}\cite{extreme_binning09}. We compared our solution to
content-based hash partitioning by running the algorithm on our set of 
traces. We used 2MB superchunks and 4KB chunks to partition the data, and use
the minhashes of the superchunks for routing. The results of this are shown in
Table~\ref{??}. Routing each chunk to a bin provides good deduplication
efficiency, and only requires each storage node to keep an index of local data
and the bin mapping, but misses significant opportunities in our indended use
case.

The problem with the Data Routing algorithm is in the very high network traffic.
Our intended use case is a backup system which runs alongside a number of VMs,
in order to save costs. Data Routing makes no use of the inherent locality in
such a system, and therefore puts a much higher network burden on machines which
are also running 35 VMs each. This will reduce the available network bandwidth
to users of the VMs, and/or reduce the possible deduplication throughput.