### 2013-06-26 snapshot_system/snapshot.tex

\section{VM-centric Snapshot Deduplication}
\label{sect:dedupe}
%snapshot representation

%We use the The deduplication process is first conducted among snapshots within each VM
%and then is conducted across VMs.
%Given the concern that searching duplicates across VMs is a global feature which can affect parallel performance
%and complicate failure management,
%we only eliminate the duplication of a small but popular data set while still maintaining a cost-effective deduplication ratio.
%For this purpose, we exploit the data characteristics of snapshots and collect most popular data.
%Data sharing across VMs is limited within this small data set such that adding replicas for it could enhance fault tolerance

%A deduplication scheme compares the fingerprints of the current snapshot
%with its parent snapshot and also other snapshots in the entire cloud.
%The traditional approach compares all fingerprints and 
%stores one copy for all of a chunk's duplicates.
%We call this as  the VM-oblivious (VO) approach.
Our VC design has the following objectives:
\begin{itemize}

\item
1) Localize the deduplication and data blocking within each VM as much as possible so 
that any failure of a file system block mainly affects the associated VM.
Localizing the snapshot data deduplication within a VM
improves the system by increasing data independence between different VM backups,
simplifying snapshot management and statistics collection,
and facilitating parallel execution of snapshot operations.
\item
2) Maintain a competitive deduplication effiency while being resource-friendly to other
existing cloud services.
Cross-VM duplication can be desirable since  there are many cloud images
use widely-used software and libraries and their data blocks are duplicate. 
As the result, different VMs tend to backup large amount of highly similar data.
\item
3)  Mininimize the number of data chunks shared among VMs and add extra replicas to incrase
fault resilience for such data chunks.
\end{itemize}

\subsection{Key VC Strategies}

%Inner-VM duplication can be very effective within VM's snapshots. There
%are typically a large number of duplicates existing among the snapshots
%for a single VM.

%Our multi-level pipeline process can minimize 
%the cost of deduplication while maximize the its efficiency at each level,
%and it is parallel since each segment is processed independently.

%Our VC differentiates duplicates within a VM and cross VMs,
%and conducts \textit{Inner-VM} and \textit{cross-VM} detection separately.

\begin{itemize}
\item \textbf{Local duplicate search.}
We start with the standard dirty bit approach in a coarse grain segment level
and use the  Xen virtual device driver to implement a feature called ``changed block tracking''
for the storage device and the dirty bit setting is maintained in a coarse grain level we call it a segment.
In our implementation, the segment size is 2MB. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment  size.

It should be noted that dirtybit tracking is supported or can be easily implemented in many 
major virtualization solution vendors. 
%(VMware, Xen, Microsoft).  
%VMware support it directly; 
The VMWare hypervisor has an API to let external backup application know 
the changed areas since last backup. Xen doesn't directly support it, However, their open-source 
architecture allows anyone to extend the device driver, thus enabling changed block tracking. 
We implement dirty bit tracking this way in Alibaba's platform.  
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic, therefore changed block tracking can be implemented externally.

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{??},  
%This allows the system to achieve deduplication even when there are insertions/deletions which would
%affect many fixed-size chunks.
we conduct additional local inner-VM deduplication by comparing
chunk signatures within a dirty segment from its parent snapshot. 
We load the chunk fingerprints in the corresponding segment from the
parent and perform fingerprint matching for further inner-VM deduplication.
The amount of memory for maintaining those fingerprints  is small, as we only
load one segment at a time.
For example, with a 2MB segment, there are about 500 fingerprints to compare.

%If we use 4KB in level-1, then such a level-1 should have similar dedup efficiency 
%as the current level-1 and level-2 combined, because finally they equal to comparing  
%parent snapshot at 4KB granularity.
%
%However, at level-3 things are different. If we use 4KB fix-size block uniformly, it would be harder for different VM to share data through CDS, because there is no guarantee that the location of duplicate data on different VM disks are always aligned at 4KB boundary. For example, if two VMs each has a copy of duplicate data, but they are not aligned, then we won't be able to detect them. Our study at the current small data set has shown that using 4KB fix-size block will make CDS method less efficient by nearly 10%. Over the long time, more and more OS variations will co-exist in the cluster, making this 4KB fix-size approach inefficient in reducing duplicate data across VMs.

%\item \textit{Level 2  Chunk fingerprint comparison.}
%If a segment is modified, we perform fine-grained deduplication 
%by comparing the fingerprints of its chunks to the same segment's recipe in the previous snapshot,
%thus eliminate partial data duplication within the segment.
%\end{itemize}
%
%In general, operations at level 1 have almost no cost and most of unmodified data are filtered here. 
%To process a dirty segment at level 2, 
%there requires no more than one DFS access to load the segment recipe from previous snapshot,
%and a tiny amount of memory to hold it in main memory.

\item \textbf{Cross-VM deduplication with top $k$ popular chunks}
This step accomplishes the standard global fingerprint  comparison as conducted
in the previous work~\cite{??}.
One key observation is that the inner deduplication has removed many of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the most popular items. 
%Since duplicate sharing patterns among  VM follows a zip-like distribution,  
\end{itemize}

When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunk. We manage this accumulation process using an append-store  scheme
and discuss this in details in Section~\ref{sect:arch}.
The system allows all machines conduct the backup in parallel in different machines, and each machine
conducts the backup of one VM at a time, and thus only requires a write  buffer for one VM.

It should be noted that CDS chunks are computed periodically and all CDS chunks are stored in a separate
append-store instance. In this way, each file block for non-CDS chunks is associated with one VM
and does not contain any CDS chunk.

 \subsection{Popular Chunk Management}
%VM-oriented Fault Isolation}
%Our objective for fault isolation is to minimize the number of VMs affected when there are failures
%in the cluster.  
%The inner-VM deduplication does not depend on any global service and the comparison
%for each VM is localized within the parent and the current snapshot.
%Lack of data dependence between VMs brings 
With cross-VM deduplication, shared data chunks
create an artificial dependence  among VMs. 
Thus we only maintain the index for the most popular chunks to reduce the inter-VM dependence
while still accomplishing competitive deduplication efficiency.

The management for popular data chunks contains two aspects.
\begin{itemize}
\item Compute and select top-$k$ most popular chunks.
The popularity of a chunk is the number  of data chunks  from different VMs
that are duplicates of this chunk after the inner VM deduplication.
This number can be computed periodically in a weekly basis.
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunks.  For cross-VM comparison,
we only store  top $k$ items and $k$ is chosen to be relatively small.
Our analysis below will show that the algorithm can still deliver competitive
deduplication efficiency after making these approximations.
\item  Since $k$ is small and these top $k$ chunks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular chunks to enhance the fault resilience.
We will provide an analysis of the space need and fault tolerance in the next subsection.
\end{itemize}

This section analyzes how value   $k$ impacts the deduplication efficiency.
%It also  compares the  fault resilience of our VM-centric deduplication approach with a standand approach using 
%global deduplication.
%\subsubsection{Impact of CDS deduplication}
The analysis is based on VM traces collected
using Alibaba's production user data~\cite{ieeecloud} 
which shows that the popularity of data chunks after inner VM deduplication follows 
a Zipf-like distribution\cite{zipf} and its
exponent $\alpha$ is ranged between 0.65  and  0.70.
%As a result, it can be proved that deduplication efficiency of CDS index is scalable:

\begin{figure}
\centering
%\epsfig{file=log-log.disk.pdf,width=3in}
%source ../vm_snapshot_sim/images/log-log.disk.eps
 \includegraphics[width=3in]{log-log-disk.pdf}
\caption{Duplicate frequency versus  chunk ranking in a log scale.}
\label{fig:Datazipf}
\end{figure}

\begin{table}[ht]
\centering
\begin{tabular}{|p{1.25cm}|p{6.5cm}|}
\hline
$b$ &  the total amount of data chunks\\ 
\hline
$b_u$ &  the total amount of unique fingerprints after inner VM  deduplication\\
\hline
$C_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$\delta$ &  the percentage of duplicates detected in inner VM deduplication\\
\hline
$\sigma$ &  the number of unique non-CDS chunks over  the number of the CDS chunks.\\
\hline
$p$ & the number of machines in the cluster\\
\hline
$D$ & the amount of unique data on each machine\\
\hline
$B$ & the average data chunk size. Our setting is  4K.\\
\hline
$s$ & the average size of file system blocks in the distributed file system. The default is  64MB.\\
\hline
$m$ & memory size on each node used by VC\\ 
\hline
$F$ & the size of an popular data index entry\\
$N_1$ & the average number  of non-CDS file system blocks  in a VM\\
$N_2$ & the average number  of CDS file system blocks  in a VM\\
$N_o$ & the average number  of file system blocks  in a VM for VO\\
\hline
\end{tabular}
\caption{Modeling  symbols.}
\label{tab:symbol}
\end{table}

{\it [Need to find a place to put these numbers in: Total number of chunks in 350 snapshots: 1,546,635,485. Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
Let $b$ be the total number of data chunks after localized deduplication, 
$b_u$ be the total number of fingerprints 
in the global index after complete deduplication, and
$C_i$ be the frequency for the $i$th most popular fingerprint. 
By Zipf-like distribution, $C_i = \frac{C_1}{i^\alpha}.$
Since $ \sum_{i=1}^{b_u}C_i = b$,
\[
C_1 \sum_{i=1}^{b_u}\frac{1}{i^\alpha} = b
\]
Given $\alpha <1$, $C_1$ can be approximated with integration:
\begin{equation}
C_1=\frac{b(1-\alpha)}{b_u^{1-\alpha}}.
\end{equation}

%We define deduplication efficiency $e$ to be the proportion of raw data that can be deduplicated,
%In complete deduplication scenario, it deduplication efficiency $e_c$ is:
%\begin{equation}
%e_c = \frac{b-b_u}{b} = 1 - \frac{b_u}{b}
%\end{equation}

The  $k$ most popular fingerprints can cover the following number of chunks after inner VM 
deduplication.
\[
C_1 \sum_{i=1}^{k}\frac{1}{i^\alpha} \approx  
C_1 \int_{1}^{k}\frac{1}{x^\alpha} dx  \approx  C_1\frac{  k^{1-\alpha}} {1-\alpha}.
\]

Deduplication efficiency of VC using top $k$ popular chunks
is the percentage of duplicates that can be detected:  
\begin{equation}
\begin{split}
e_k &= 
\frac{ b(1-\delta) + C_1\frac{  k^{1-\alpha}} {1-\alpha}} 
{b(1-\delta)  + \delta b - b_u }\\
&= 
\frac{ (1-\delta) + \delta  (\frac{k}{b_u})^{1-\alpha}}
{ 1- \frac{b_u}{b} }.
\end{split}
\end{equation}

%\begin{equation}
%\begin{split}
%e_k &= (\sum_{i=1}^{k}C_i - k)/{b} = (b * \frac{1-\alpha}{b_u^{1-\alpha}} * \sum_{i=1}^{k}\frac{1}{i^\alpha} - k)/b \\
%&\approx \frac{b*\frac{1-\alpha}{b_u^{1-\alpha}}*\frac{k^{1-alpha}}{1-\alpha} - k}{b}
%&\approx (\frac{k}{b_u})^{1-\alpha} - \frac{k}{b} \approx (\frac{k}{b_u})^{1-\alpha},\; (since\; k \ll b)
%\end{split}
%\end{equation}

% For the Zipf-like distribution, an approximation to the sum of the first $n$ 
% elements of the distribution can be derived as follows:
% \begin{equation}
% \sum_{i=1}^{n}\frac{1}{i^\alpha}\approx \int_{1}^{n}\frac{1}{x^\alpha}\mathrm{d}x=\frac{x^{1-\alpha}}{1-\alpha}=\frac{n^{1-\alpha}}{1-\alpha}\;  for\;  \alpha<1
% \end{equation}

% So the cumulative distribution function for a CDS holding top $S_c$ fingerprints
% of global index with size $S_g$ is:

% \begin{equation}
%   E = (S_c / S_g)^{1-\alpha} \;  for\;  \alpha<1
% \end{equation}

%[Describe the dedup efficiency model in detail]
Let $p$ be the number of physical machines in the cluster, $m$ be the memory on each node used by the popular
index, $F$ be the size of an index entry,
$D$ be the amount of unique data on each physical machine, and 
$B$ be the average chunks size. We store the popular index using a distributed shared memory hashtable
such as MemCacheD.  Then $k$ and $b_u$ can be expressed as:
$
k = p*m/F$, and $b_u = p*D/B.
$

The overall deduplication efficiency of VC is
\[
\frac{ (1-\delta) + \delta (\frac{m*B}{D*F})^{1-\alpha}}
{ 1- \frac{b_u}{b} }.
\]
where $(\frac{m*B}{D*F})^{1-\alpha}$ represents the percentage of the remaining chunks detected as duplicates
after inner VM deduplication. 
%Then the deduplication efficiency of CDS approach using top k most popular fingerprints is:
%\begin{equation}
%  e_k = (\frac{m*B}{F*D})^{1-\alpha}
%\end{equation}

When the number of machines at each cluster increases, the number of total VMs increases.
Then $k$ increases since more memory is available to host the popular chunks index.
But for each physical machine, the number of VMs remains the same, and thus
$D$ is  a constant. Then  the overall deduplication efficiency of VC remains
a constant.

% is mainly affected by the its memory usage,
%and the Zipf-like distribution parameter $\alpha$.
%From the memory usage perspective, the more memory we give to CDS index, the more duplicates it can found.
%From the data characteristics perspective, if the system tends to backup arbitrarily random data, 
%$\alpha$ would be small and thus CDS approach would not be effective. But in a VM cluster all VMs
%have a large portion of similar data, thus leading to a bigger $\alpha$. The more VM join the cluster,
%the better deduplication efficiency our CDS deduplication approach will have.

\comments{
Since $B$, $D$ and $F$ are environment constants, the deduplication efficiency of CDS is mainly affected by its memory usage,
and the Zipf-like distribution parameter $\alpha$.
From the memory usage perspective, the more memory we give to CDS index, the more duplicates it can found.
From the data characteristics perspective, if the system tends to backup arbitrarily random data, 
$\alpha$ would be small and thus CDS approach would not be effective. But in a VM cluster all VMs
have a large portion of similar data, thus leading to a bigger $\alpha$. The more VM join the cluster,
the better deduplication efficiency our CDS deduplication approach will have.
}

\begin{table}
    \begin{tabular}{llll}
    Data size (GB) & 1\%    & 2\%    & 4\%    \\
    14.6           & 18.6\% & 22.1\% & 31.4\% \\
    28.1           & 19.5\% & 26.2\% & 38.8\% \\
    44.2           & 21.7\% & 26.5\% & 36\%   \\
    61.6           & 23.2\% & 32.9\% & 35\%   \\
    74.2           & 23.6\% & 33.6\% & 37.5\% \\
    \end{tabular}
    \caption{Deduplication effectiveness of top k\% of global index}
    \label{tab:cds}
\end{table}

\begin{figure}[ht]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={CDS Coverage},
            xlabel={Total Data Stored (GB)},
            ylabel={CDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            mark options=solid,
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.2)},
            anchor=north}
            ]
            \addplot[color=blue,mark=*] table[x=InputSize,y=A1] {figures/cds_coverage.txt};
            \addplot[color=blue,mark=square*] table[x=InputSize,y=A2] {figures/cds_coverage.txt};
            \addplot[color=blue,mark=triangle*,mark size=3] table[x=InputSize,y=A4] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=*] table[x=InputSize,y=T1] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=square*] table[x=InputSize,y=T2] {figures/cds_coverage.txt};
            \addplot[densely dashed,color=red,mark=triangle*] table[x=InputSize,y=T4] {figures/cds_coverage.txt};
            \legend{Measured (1\%CDS),Measured (2\%CDS),Measured (4\%CDS),Predicted (1\%CDS),Predicted (2\%CDS),Predicted (4\%CDS)};
            \end{axis}
    \end{tikzpicture}
  \caption{fixed-alpha predicted vs. actual CDS coverage as data size increases.}
  \label{fig-vo-links}
\end{figure}