### 2013-07-11 snapshot_system/main.tex

\documentclass{sig-alternate-10pt}
\usepackage{mathptmx}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{array}
\usepackage{breqn}
\usepackage{pgfplots}
\pgfplotsset{compat=1.7}
\DeclareGraphicsExtensions{.pdf}

\newcommand{\comments}[1]{}
%pgfplots setup
\pgfplotsset{every axis/.append style={
	very thick,
	tick style={thick}}}
\pgfplotsset{
        every mark/.append style={solid}
}
\pgfplotscreateplotcyclelist{mcolor}{%
    {blue,every mark/.append style={fill=blue},mark=*},
    {red,every mark/.append style={fill=red},mark=square*},
    {brown,every mark/.append style={fill=brown},mark=triangle*,mark size=3},
    {black,every mark/.append style={fill=black},mark=diamond*,mark size=3},
    {green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=2.5},
    {magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=3},
    {densely dashed,blue,every mark/.append style={fill=blue},mark=*},
    {densely dashed,red,every mark/.append style={fill=red},mark=square*},
    {densely dashed,brown,every mark/.append style={fill=brown},mark=triangle*,mark size=2.7},
    {densely dashed,black,every mark/.append style={fill=black},mark=diamond*,mark size=2.7},
    {densely dashed,green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=2.5},
    {densely dashed,magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=3}}

\pgfplotscreateplotcyclelist{mmark*}{
    {mark=*},
    {mark=square*},
    {mark=triangle*,mark size=2.7},
    {mark=diamond*,mark size=2.7},
    {mark=pentagon*,mark size=2.5},
    {mark=star,mark size=3},
    {mark=asterisk,mark size=3},
    {mark=+,mark size=3},
    {mark=Mercedes star,mark size=3}}

\pgfplotscreateplotcyclelist{mline}{
    {solid,blue,every mark/.append style={fill=blue}},
    {densely dashed,red,every mark/.append style={fill=red}},
    {densely dotted,brown,every mark/.append style={fill=brown}},
    {loosely dotted,black,every mark/.append style={fill=black}},
    {loosely dashed,green!60!black,every mark/.append style={fill=green!60!black}},
    {densely dashdotted,magenta!90!black,every mark/.append style={fill=magenta!90!black}},
    {densely dashdotdotted,teal,every mark/.append style={fill=teal}},
    {loosely dashdotted,orange,every mark/.append style={fill=orange}},
    {loosely dashdotdotted,violet,every mark/.append style={fill=violet}}}

    \pgfplotsset{cycle list name=mcolor}

\begin{document}

%\title{Deduplication Localization for Virtual Machine Snapshot Backup}
\title{Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup}
%\title{BigArchive: Low-cost and Scalable Deduplication Storage for VM Snapshot}

\author{
Wei Zhang, Michael Agun,  Tao Yang \\
%Wei Zhang$^{\star}$, Michael Agun,  Tao Yang$^\star$
%{\normalsize$^\star$  University of California at Santa Barbara}, {\normalsize$^\dagger$Alibaba Inc.}
%Wei Zhang, Tao Yang, Gautham Narayanasamy\\
Department of Computer Science\\
University of California,  Santa Barbara, CA 93106
%\and
%Hong Tang\\
%Alibaba Inc.
}
%, Gautham Narayanasamy$^\star$, and Hong Tang$^\dagger$ \\

\maketitle

\input{abs}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Theory}

%\keywords{ACM proceedings, \LaTeX, text tagging} % NOT required for Proceedings

\input{intro}
%\input{design}
\input{background}
\input{challenge}
\input{snapshot}
\input{failover}
%\input{compare}
%\input{design1}
\input{arch}
%\input{impl}
%\input{sampling-comparison.txt}
\input{eval1}
%\input{cds}
%\input{delete}
%\input{failover}
%\input{result}
\input{con}

\bibliographystyle{abbrv}
\bibliography{library}
\end{document}