### 2013-07-16 snapshot_system/cds.tex

\section{PDS Analysis}
[Purpose of using PDS]
PDS is a small percentage of the global data index that we into DHT as a shared index. 
With PDS as as a public data resource, each VM's snapshot backup has no need to 
store its own copies for data that already exist in PDS, instead they can just
store a reference. 
We expect PDS to contain most of the data related to operating system, software and libraries.

[What's the benifits]
Using PDS as a common data source brings us several benefits: 
First, the data we collect into PDS are commonly used in many VMs, 
thus the PDS index, a special subset of global index, has very high deduplication efficency.
Second, in our architecture the PDS data are the only data that can be shared between different VMs,
by carefully control the replication degree of PDS data, we could improve the fault-tolerancy
in such a data sharing environment. Finally, the PDS index is small enough to be put into a 
in-memory DHT, thus it avoids the excess memory usage of global index 
while brings us the majority of deduplication efficiency of it.

[Describe how PDS is generated]

[Explain why PDS scales effectively as data grow]
Our empirical study based on VM images from production environment\cite{ieeecloud} showed that the 
frequency of data duplication follows Zipf-like distribution\cite{zipf}, 
with the exponent $\alpha$ between 0.65 ~ 0.70. 
As a result, it can be proved that deduplication efficiency of PDS index is scalable:

[Describe the dedup efficiency model in detail]
Let $S_c$ be the total number of FPs in PDS, $S_g$ be the total number of FPs in global index, $F$ be the size of a FP, $N$ be the number of nodes in the cluster, $m$ be the memory on each node that are used by PDS, $D$ be the amount of data on each node, and $B$ be the average block size. The deduplication efficiency will be:
\begin{equation}
  E = (S_c / S_g)^{1-\alpha}
\end{equation}
The $S_c$ and $S_g$ can be expressed as:
\begin{equation}
S_c = N*m/F, S_g = N*D/B
\end{equation}
By replacing $S_c$ and $S_g$ in the first formula, we have:
\begin{equation}
  E = (\frac{m*B}{F*D})^{1-\alpha}
\end{equation}

Since $B$, $D$ and $F$ are pre-configured constants, the deduplication efficiency of PDS is only controlled by the its memory usage.