### 2013-07-26 batchdedup/main.tex

\documentclass[finalversion]{usetex-v1}
%\usepackage[top=1in, bottom=1in, left=1.25in, right=1.25in]{geometry}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{epsfig}
%\usepackage{breakurl}
%% Define a new 'leo' style for the package that will use a smaller font.
%\makeatletter
%\def\url@leostyle{%
%  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\ttfamily}}}
%\makeatother
%% Now actually use the newly defined style.
%\urlstyle{leo}
%\linespread{1.2}
%\setlength{\parskip}{1ex}

\usepackage{pgfplots}
\pgfplotsset{compat=1.7}
\DeclareGraphicsExtensions{.pdf}

%pgfplots setup
\pgfplotsset{every axis/.append style={
	very thick,
	tick style={thick}}}
\pgfplotsset{
        every mark/.append style={solid}
}
\pgfplotscreateplotcyclelist{mcolor}{%
    {blue,every mark/.append style={fill=blue},mark=*},
    {red,every mark/.append style={fill=red},mark=square*},
    {brown,every mark/.append style={fill=brown},mark=triangle*,mark size=3},
    {black,every mark/.append style={fill=black},mark=diamond*,mark size=3},
    {green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=2.5},
    {magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=3},
    {densely dashed,blue,every mark/.append style={fill=blue},mark=*},
    {densely dashed,red,every mark/.append style={fill=red},mark=square*},
    {densely dashed,brown,every mark/.append style={fill=brown},mark=triangle*,mark size=2.7},
    {densely dashed,black,every mark/.append style={fill=black},mark=diamond*,mark size=2.7},
    {densely dashed,green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=2.5},
    {densely dashed,magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=3}}

\pgfplotscreateplotcyclelist{mmark*}{
    {mark=*},
    {mark=square*},
    {mark=triangle*,mark size=2.7},
    {mark=diamond*,mark size=2.7},
    {mark=pentagon*,mark size=2.5},
    {mark=star,mark size=3},
    {mark=asterisk,mark size=3},
    {mark=+,mark size=3},
    {mark=Mercedes star,mark size=3}}

\pgfplotscreateplotcyclelist{mline}{
    {solid,blue,every mark/.append style={fill=blue}},
    {densely dashed,red,every mark/.append style={fill=red}},
    {densely dotted,brown,every mark/.append style={fill=brown}},
    {loosely dotted,black,every mark/.append style={fill=black}},
    {loosely dashed,green!60!black,every mark/.append style={fill=green!60!black}},
    {densely dashdotted,magenta!90!black,every mark/.append style={fill=magenta!90!black}},
    {densely dashdotdotted,teal,every mark/.append style={fill=teal}},
    {loosely dashdotted,orange,every mark/.append style={fill=orange}},
    {loosely dashdotdotted,violet,every mark/.append style={fill=violet}}}

    \pgfplotsset{cycle list name=mcolor}

\newcommand{\comments}[1]{}

\begin{document}
%\title{Selective Deduplication for VM Snapshots in Cloud Storage}
\title{ Low-Cost Data Deduplication  for  Virtual Machine Backup in Cloud Storage}
%\title{Data Deduplication and Zipf-like Distribution in VM Cloud Storage}

\author{
Wei Zhang$^{\star}$, Tao Yang$^\star$, Gautham Narayanasamy$^\star$, and Hong Tang$^\dagger$ \\
{\normalsize$^\star$  University of California at Santa Barbara}, {\normalsize$^\dagger$Alibaba Inc.}
%Wei Zhang, Tao Yang, Gautham Narayanasamy\\
%University of California at Santa Barbara
%\and
%Hong Tang\\
%Alibaba Inc.
}

% \{wei, tyang\}@cs.ucsb.edu

\date{}
\maketitle

%need to explain: vm cloud, vm backup, dedup fundamentals, vm dedup(2 works), distributed dedup, zipf
%need to define: requirements, design considerations, architecture

\input{abs}
\input{intro} %introduce the vm cloud, backup problem, and our requirements
\input{background} %related works, related solutions, 
\input{arch} %, design considerations, architecture
%\input{analysis} %, design considerations, architecture
\input{eval} %data study, zipf, model of cds size and dedup efficiency
%\input{results}
%\input{related}
%\input{future}
\input{concl}

{\bf Acknowledgment.} We thank Michael Agun, Renu Tewari, and
the anonymous referees for their valuable comments.
This work is supported in part by NSF IIS-1118106. 
Any opinions, findings, conclusions or recommendations expressed in this material
are those of the authors and
do not necessarily reflect the views of the National Science Foundation.

      %\setlength{\itemsep}{ex}%
%      \setlength{\parskip}{0ex}%
%\setlength{\itemsep}{-3mm}

	%\linespread{0.3} 
%  \let\oldthebibliography=\thebibliography
%  \let\endoldthebibliography=\endthebibliography
%  \renewenvironment{thebibliography}[1]{%
%    \begin{oldthebibliography}{#1}%
%      \setlength{\parskip}{-0.02ex}%
%      \setlength{\itemsep}{-0.02ex}%
%      \setlength{\baselineskip}{-0.02ex}%

%  }%
%  {%
%    \end{oldthebibliography}%
%  }
{\small
\bibliographystyle{abbrv}
\bibliography{library,dedup1}
}
\end{document}