### 2013-07-26 snapshot_system/abs.tex

\begin{abstract}
A cloud environment that  hosts a large number of  virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots and
signature-based deduplication of data blocks is necessary to eliminate excessive 
redundant blocks.  Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
% when their signatures are identical. 
however it is resource expensive and less fault-resilient to perform a global deduplication
and let a data block share by many virtual machines. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine
and associates  underlying file blocks with one VM for most of cases.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resource. 
This  paper  describes a comparative evaluation of this scheme in  accomplishing a high deduplication 
efficiency while sustaining a good backup throughput.

%This paper studies the      a VM snapshot storage architecture which adopts multiple-level selective deduplication to bring the benefits of fine-grained data reduction into cloud backup storage systems.
%In this work, we describe our working snapshot system implementation, and provide
%early performance measurements for both deduplication impact and
%snapshot operations.
\end{abstract}