### 2013-08-25 snapshot_system/abs.tex
revise the analysis

\begin{abstract}
A cloud environment that  hosts a large number of  virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots and
signature-based deduplication of data blocks is necessary to eliminate excessive 
redundant blocks.  Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
% when their signatures are identical. 
however it is resource expensive and less fault-resilient to perform a global deduplication
and let a data block share by many virtual machines. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine using similarity-guided
search and associates  underlying file blocks with one VM for most of cases.
It restricts cross-VM deduplication under common blocks with extra replication support.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resource. 
This  paper  describes a comparative evaluation of this scheme in  accomplishing a competitive deduplication 
efficiency while sustaining a good backup throughput.

%This paper studies the      a VM snapshot storage architecture which adopts multiple-level selective deduplication to bring the benefits of fine-grained data reduction into cloud backup storage systems.
%In this work, we describe our working snapshot system implementation, and provide
%early performance measurements for both deduplication impact and
%snapshot operations.
\end{abstract}