### 2013-08-28 snapshot_system/concl.tex
revise eval1.tex

\section{Conclusion}
\label{sect:final}
In this paper we propose a collocated backup service built on
the top of a cloud cluster to reduce network traffic and infrastructure.
The key contribution is a VM-centric deduplication scheme to 
maximize fault isolation while delivering competitive deduplication efficiency.
Inner-VM deduplication localizes backup data dependency and exposes more parallelism  
while cross-VM deduplication with a small common data set.
VM-specific file block packing also reduces inter-VM dependence and enhances fault tolerance.
The design places a special consideration for low-resource usages as a collocated cloud service.
Evaluation using  VM backup data shows that
our solution can accomplish 92\% of what complete global
deduplication can do while the availability of snapshots increases substantially with this VM-centric
and a small replication overhead for popular inter-VM chunks.

% effectively  covers a large percent of duplicate data.
%The scheme organizes the write of small data chunks into large file system blocks so
%that each underlying file block is associated with one VM for most cases.
%by keeping most of the FSBs only referenced by 1 VM, we isolate faults and improve overal fault tolerance.
%Our solution accomplishes reasonable and competitive deduplication efficiency while
%still meeting stringent cloud resource usage requirements. 
%Our analysis shows that  this VM centric scheme 
%can  provide  better fault tolerance than VM-oblivious global deduplication schemes
%while using a small amount of computing and storage resources.

%[Talk about more what we learn from Evaluation]
%Compare to today's widely-used snapshot technique, our scheme reduces almost
%two-third of snapshot storage cost.
%Finally, our scheme uses a very small amount of memory on each node, and leaves
%room for additional optimization we are further studying.

%This paper studies a cluster-based VM-centric scheme which collocates a lightweight
%backup service with the cloud service and it integrates multiple duplicate detection strategies that  
%localize the deduplication as much as possible within each virtual machine.
%This  paper  provides a comparative evaluation of this scheme in  accomplishing a high deduplication 
%efficiency while sustaining a good backup throughput.

%This paper studies the      a VM snapshot storage architecture which adopts multiple-level selective deduplication to bring the benefits of fine-grained data reduction into cloud backup storage systems.
%In this work, we describe our working snapshot system implementation, and provide
%early performance measurements for both deduplication impact and
%snapshot operations.

\comments{
\section*{Acknowledgments}
{
We would like to thank Weicai Chen and Shikun Tian from Alibaba for their kind support, 
and the anonymous referees for their comments.
Wei Zhang has received internship support from Alibaba  for VM backup system development.
This work was supported in part by NSF IIS-1118106.
Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and
do not necessarily reflect the views of Alibaba or the National Science Foundation.
}
}
%Noted that 6\% is still significant, which is about 24GB per each VM and for a 1000 node Aliyun cluster,
%this is about 600 terabytes.

%Our experiments show th
%our solution can eliminate the majority of data duplication with a tiny fraction of
%block hash index store in memory. It does not only saves valuable system resouces in
%the VM cloud, but also makes deduplication much faster.
%
%
%Using  50 user VM data out of 1322 data disks as the training data and
%with  1.5\% as PDS threshold, we see the total 1198GB of new data is reduced by
%755.8GB, while perfect deduplication can reduce 1017.4GB. So 74.3\% of duplicate blocks are eliminated
%by pre-trained PDS, which is quite satisfactory.