### 2013-09-03 snapshot_system/fast14_conclusion.tex
fix citations

\section{Conclusion}
\label{sect:conclusion}
In this paper we propose a collocated backup service built on
the top of a cloud cluster to reduce network traffic and infrastructure.
The key contribution is a VM-centric deduplication scheme to 
maximize fault isolation while delivering competitive deduplication efficiency.
Inner-VM deduplication localizes backup data dependency and exposes more parallelism  
while cross-VM deduplication with a small common data set.
VM-specific file block packing also reduces inter-VM dependence and enhances fault tolerance.
The design places a special consideration for low-resource usages as a collocated cloud service.
Evaluation using  VM backup data shows that
our solution can accomplish 92\% of what complete global
deduplication can do while the availability of snapshots increases substantially with this VM-centric
and a small replication overhead for popular inter-VM chunks.