### 2013-09-13 snapshot_system/impl.tex

\section{Implementation}
\subsection{Snapshot Backup}
\subsection{Snapshot Read}
\subsection{Snapshot Deletion}
The following steps would take place during an approximate deletion:

\begin{enumerate}
\item {\bf Creating Bloom filter} Scan all the living snapshot recipess and their segment recipes,
for every reference pointing to append store, add it to the Bloom filter.
\item {\bf Check existance} For every data reference in the deleted snapshot recipe and its segment recipes,
check the existance of that data reference in Bloom filter. If not found, it is safe to delete that piece of data from append store
because no living snapshots has referenced it.
\end{enumerate}

The overall time of running a approximate deletion for one snapshot deletion would be scanning
all the living snapshots and deleted snapshots, since operations on the in-memory Bloom filter can be done in
parallel and is much faster than loading recipes from DFS:
\begin{equation}
T = (N_{SS} + 1) * T_{scan\_recipes}
\end{equation}

Using the example and analysis in previous section, this approximate deletion can be done in 5 minutes. 
Memory usage of the Bloom filter depends on its false-positive probility $P_{bl}$,
when set $P_{bl}$ to 0.01, the memory footprint of approximate deletion is about 15 MB.