### 2013-09-19 snapshot_system/fast14_deduplication.tex

\section{VM-centric Snapshot Deduplication}
\label{sect:deduplication}

\subsection{Key VM-centric  Strategies}
\label{sect:vc-strategies}
\begin{itemize}
\item \textbf{VM-specific local duplicate search within similar segments.}
We start with the standard dirty bit approach in a coarse grain segment level.
In our implementation with Xen on an Alibaba platform, the segment size is 2MB
and  the device driver is extended to support  changed block tracking. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment  size.
It should be noted that dirtybit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement  the changed block tracking feature.

\begin{figure}[htbp]
  \centering
  \epsfig{file=images/mh_local_dedup.png, width=3in}
  \caption{Similarity-guided local duplicate detection}
  \label{fig:local_dedup}
\end{figure}

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{Jin2009},
we conduct additional local inner-VM deduplication on a snapshot by comparing
chunk fingerprints of a dirty segment 
with those in  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their content signature is the same.
This segment content signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). Note that this definition of
content similarity is  an approximation~\cite{AdnriBroader}.  When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot recipe.  Then recipes of the similar segments are loaded to memory,
which contain chunk fingerpints which can  potentially be duplicates.
To limit the search scope, we set a limit on the the number of  similar segments to be  searched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For example, a 2MB segment contains about 500 fingerprints and by limiting at most 10 similar
segments to search, the amount of memory for maintaining those fingerprints in similar segments is small.

\item \textbf{Cross-VM deduplication with popular chunks and replication support}
This step accomplishes the standard global fingerprint comparison as conducted
in the previous work~\cite{WeiZhangIEEE}.
One key observation is that the inner-VM deduplication has removed many of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the top $k$
most popular items. This dataset is called the PDS (popular data set). 
The popularity of a chunk is the number  of data chunks  from different VMs
that are duplicates of this chunk after the inner VM deduplication.
This number can be computed periodically (e.g., on a weekly basis).
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk signatures in a distributed shared memory.

Since $k$ is relatively small and these top $k$ chunks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular chunks to enhance the fault resilience.

\item \textbf{VM-centric file system block management.}
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. We manage this accumulation process using an append-store  scheme built
on a distributed file system discussed  in Section~\ref{sect:architecture}.
%Non-PDS chunks from the same VM are stored in one append store. 
Each file system block is either dedicated to non-PDS chunks, or PDS-chunks.
A file system block for non-PDS chunks is associated with one VM and does not contain
any PDS chunks, to maintain our goal of fault isolation.
PDS chunks are stored in a separate append-store instance with special replication handling. 
The system allows all machines conduct the backup in parallel, and each machine
conducts the backup of one VM at a time, and thus only requires a write  buffer for one VM.
\end{itemize}

\begin{figure}
\centering
%\epsfig{file=log-log.disk,width=3in}
%source ../vm_snapshot_sim/images/log-log.disk.eps
 \includegraphics[width=3in]{figures/log-log-disk}
\caption{Duplicate frequency versus  chunk ranking in a log scale after inner-VM deduplication.}
\label{fig:Datazipf}
\end{figure}
We have not  adopted the  sampled index~\cite{Guo2011} 
for  detecting duplicates from popular chunks.
For the data sets we have tested, the spatial locality 
among popular
data chunks is limited and  on average the number of consecutive data chunks is 7 among popular chunks.
This is not sufficient to enable prefetching which is required  for index sampling.

\begin{table}[htbp]
\centering
\begin{tabular}{|p{1.25cm}|p{5.5cm}|}
\hline
$k$ &  the number of top most popular chunks selected for deduplication\\ 
\hline
$c$ &  the total amount of data chunks in a cluster of VMs\\ 
\hline
$c_u$ &  the total amount of unique fingerprints after perfect  deduplication\\
\hline
$f_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$\delta$ &  the percentage of duplicates detected in inner VM deduplication\\
\hline
$\sigma$ & =$\frac{k}{c_u}$ which is  the percentage of unique data  belong to  PDS\\
\hline
$p$ & the number of machines in the cluster\\
\hline
$V$ & the average number of VMs hosted on each machine\\
%\hline
%$D$ & the amount of unique data on each machine\\
%\hline
%$C$ & the average data chunk size. Our setting is  4K.\\
%\hline
%$s$ & the average size of file system blocks in the distributed file system. The default is  64MB.\\
%\hline
%$m$ & memory size on each node used by VC\\ 
%\hline
%$E$ & the size of an popular data index entry\\
\hline
$N_1$ & the average number  of non-PDS file system blocks  in a VM for VC\\
\hline
$N_2$ & the average number  of PDS file system blocks  in a VM for VC\\
\hline
$N_o$ & the average number  of file system blocks  in a VM for VO\\
\hline
A(r) & the availability of a file system block with replication degree $r$\\
\hline
\end{tabular}
\caption{Modeling  parameters}
\label{tab:symbol}
\end{table}

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
\subsection{Impact on deduplication efficiency}
Choosing the value  $k$ for the most popular chunks affects the deduplication efficiency.
We analyze this impact based on the characteristics  of the VM snapshot traces
studied from  application datasets.
Our previous study shows that the popularity of data chunks after inner VM deduplication follows 
a Zipf-like distribution\cite{Breslau1999a} and its
exponent $\alpha$ is ranged between 0.65  and  0.7~\cite{WeiZhangIEEE}. 
Figure~\ref{fig:Datazipf} illustrates the Zipf-like distribution of chunk popularity.
We will use a number of parameters in our analysis below and they are defined in
Table~\ref{tab:symbol}.

%Table~\ref{tab:symbol} defines paramters $c$, $c_u$, $f_i$, and $\delta$ used below.
%let $c$ be the total number of data chunks. 
%$c_u$ be the total number of fingerprints 
%in the global index after complete deduplication, and
%$f_i$ be the frequency for the $i$th most popular fingerprint. 
By Zipf-like distribution, $f_i = \frac{f_1}{i^\alpha}.$
Since $ \sum_{i=1}^{c_u}f_i = c (1-\delta)$,
\[
f_1 \sum_{i=1}^{c_u}\frac{1}{i^\alpha} = c (1-\delta).
\]
Given $\alpha <1$, $f_1$ can be approximated with integration:
\begin{equation}
f_1=\frac{c(1-\alpha)(1-\delta)}{c_u^{1-\alpha}}.
\end{equation}

The  $k$ most popular fingerprints can cover the following number of chunks after inner VM 
deduplication:
\[
f_1 \sum_{i=1}^{k}\frac{1}{i^\alpha} \approx  
f_1 \int_{1}^{k}\frac{1}{x^\alpha} dx  \approx  f_1\frac{  k^{1-\alpha}} {1-\alpha}
=c(1-\delta) \sigma^{1-\alpha}.
\]

Deduplication efficiency of the VC approach using top $k$ popular chunks
is the percentage of duplicates that can be detected:  
\begin{equation}
\label{eq:dedupeff}
%\begin{split}
%e_k &= 
\frac{ c\delta + c(1-\delta) \sigma^{1-\alpha}}
{c  - c_u }.\\
%\end{split}
\end{equation}

We store the popular index using a distributed shared memory hashtable such as MemCacheD
and allocate a fixed percentage of memory space   per physical machine for top $k$ popular items.
As the number of physical machines ($p$) increases,
the entire cloud cluster can host more VMs; however,  ratio $\sigma$ which is $k/c_u$ remains
a constant because each physical machine on average still hosts a fixed constant number of 
VMs. Then the overall deduplication efficiency of VC defined in Formula~\ref{eq:dedupeff}
remains constant.
Thus the deduplication efficiency is stable  as $p$ increases as long as $\sigma$  is a constant.

\begin{figure}[htbp]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={PDS Coverage},
            width=\linewidth,
            height=0.6\linewidth,
            cycle multi list={
                mline\nextlist
                [3 of]mmark*\nextlist
            },
            %cycle list name=mcolor,
            xlabel={Total num. chunks stored (in billions)},
            ylabel={PDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            mark options=solid,
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.30)},
            anchor=north},
            ]
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A4] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T4] {figures/cds_coverage.txt};
            \legend{Measured ($\sigma=1\%$),Measured ($\sigma=2\%$),Measured ($\sigma=4\%$),Predicted ($\sigma=1\%$),Predicted ($\sigma=2\%$),Predicted ($\sigma=4\%$)};
            \end{axis}
    \end{tikzpicture}
  \caption{Predicted vs. actual PDS coverage as data size increases.}
  \label{fig:cds-coverage}
\end{figure}
Ratio $\sigma^{1-\alpha}$ represents the percentage of the remaining
chunks detected as duplicates in global cross-VM deduplication due to PDS.
We call this PDS coverage.
Figure~\ref{fig:cds-coverage} shows predicted PDS coverage using $\sigma^{1-\alpha}$ when $\alpha$ is fixed at
0.65 and measured PDS coverage in our test dataset.
$\sigma=2\%$ represents memory usage of approximately 100MB memory per machine for top $k$ PDS.
While the predicted value remains flat, the  real coverage actually increases as the data size
increases, which indicates that $\alpha$ increases with the data size. 
Thus in practice, the coverage of the PDS increases as more VMs are involved.

\subsection{Storage Space and Impact on Fault Tolerance}

\begin{figure}
    \centering
    \subfigure[Sharing of file blocks under VC]
    {
        \includegraphics[width=3in]{images/share_vc}
        \label{fig:share_vc}
    }
    \\
    \subfigure[Sharing of file blocks under VO]
    {
        \includegraphics[width=3in]{images/share_vo}
        \label{fig:share_vo}
    }
    \caption{Bipartite association of VMs and file blocks under (a) VC and (b) VO. }
    \label{fig:share}
\end{figure}

The replication degree of the backup storage 
is $r$ for regular file blocks and $r=3$ is a typical setting in distributed
file systems~\cite{googlefs03,hdfs10}.
In the VC approach, a special replica degree $r_c$ used for PDS blocks where $r_c>r$. 
The storage cost for VO with full deduplication is $c_u *r$ and for VC, it is
$ k*r_c  + (c_u-k)*r$. Thus The storage cost ratio of $VC$ over $VO$ is 
\[
\sigma \frac{r_c}{r} + 1-\sigma
\]
Since $\sigma$ is small (it is 2\% in our experiments),  
term $\sigma \frac{r_c}{r}$ is small in the above expression.  
Thus the impact on storage increase is very small even when we choose a large $\frac{r_c}{r}$ ratio. 
For example, $\frac{r_c}{r}=2$ or 3.

Next we  assess  the impact of losing $d$ machines 
to the the VC and VO approaches.  
A large $\frac{r_c}{r}$ ratio can have a positive impact on full availability of VM snapshot blocks.

We use file system blocks rather than a deduplication
data chunk as our unit of failure because the DFS keeps
filesystem blocks as its base unit of storage.
To compute the full availability of all snapshots of a VM, we derive
the probablity of losing a snapshot file block of a VM by
estimating the number of file system blocks per VM in each approach.
As illustrated in Figure~\ref{fig:share},
we build a bipartite graph representing the association from unique file system blocks
to their corresponding VMs in each approach. An association edge is  drawn  from a file block  to a VM 
if this block is used by this VM.

For VC, each VM has an average number of $N_1$ non-PDS file system blocks
and  has  an average of  $N_2$ PDS file system blocks. 
Each non-PDS file system block is associated with one VM
 and  we denote that PDS file system blocks are
shared by an average of $V_c$ VMs. Let $s$ be the average number of chunks per file block. Then, 
\[
V *p*N_1 *s  \approx c_u (1-\sigma)\; \mbox{ and } \; 
V *p*N_2 *s  \approx c_u \sigma *V_c
\]
For the VO approach, 
For VO, each VM has an average of $N_o$ file system blocks
and let $V_o$ be the average number of VMs shared by each file system block.
\[
V *p*N_o *s  = c_u  V_o.
\]

Since each file block (with default size $64MB$) contains many chunks (on average 4KB),
each file block contains the hot low-level chunks shared by many VMs, and it also contains
rare chunks which are not shared.  From the above equations:
\[
\frac{N_1}{N_o} \approx  \frac{1-\sigma}{V_o}.
\]

Thus $N_1$ is much smaller than $N_o$. 
When there is a failure in file blocks with replication degree $r$
and there is no failure for PDS data with more replicas,   a VM in
the VC approach has a much lower chance to lose a snapshot than VO. 
Figure~\ref{fig:fsb-links} shows the average number of VMs sharing each file block from
our test dataset.
%which shows how many VMs the failure of a single block can affect. 
For VC, we only show
the sharing degree of  PDS-oriented file blocs  because non-PDS blocks always belong to
one VM.
From data, 
we have observed that $N_1 +N_2 < N_o$ in Figure~\ref{fig:vm-links}. 
This is likely because the PDS FSBs tightly pack data used by many VMs, 
which decreases the overall number of FSBs required to backup a VM.
Also if  the backup for multiple VMs were conducted concurrently, there would be more
VMs shared  by each file block on average in VO. 
%Therefore, even when there is a loss of a PDS block, the VC approach tolerates the fault better.

\begin{figure}[htbp]
  \centering
	\begin{tikzpicture}
		\begin{axis}[
                width=\linewidth,
                height=0.6\linewidth,
		%title={VO FSB links},
		xlabel={Number of VMs},
		ylabel={Avg. Num. VMs sharing FSB},
                cycle list name=mline,
                xmin=0,
                ymin=0,
                xmax=106,
		%extra y ticks={4.5,5.5,6.5} %to show extra ticks
		legend pos=north west
		]
                \addplot table[x=VMs,y=FSBLinks] {figures/vo_fsb_links_105.txt};
		\addplot table[x=VMs,y=cdslinks] {figures/cds_links_105.txt};
                \legend{VO, PDS in VC};
		\end{axis}
	\end{tikzpicture}
  \caption{Measured average number of VMs sharing a 64MB FSB in VO and in VC.}
  \label{fig:fsb-links}
\end{figure}

\begin{figure}[htbp]
  \centering
	\begin{tikzpicture}
		\begin{axis}[
		%title={VO VM links},
                width=\linewidth,
                height=0.6\linewidth,
		xlabel={Number of VMs},
		ylabel={Avg. Num. FSBs used by VM},
                xmin=0,
                ymin=0,
                xmax=106,
		%extra y ticks={4.5,5.5,6.5} %to show extra ticks
		legend pos=south west
		]
                \addplot[blue,mark=none] table[x=VMs,y=N_O] {figures/vm_links_all.txt};
                \addplot[red,dotted,mark=none] table[x=VMs,y expr=\thisrow{N_1}+\thisrow{N_2}] {figures/vm_links_all.txt};
                \addplot[red,dashdotted] table[x=VMs,y=N_1] {figures/vm_links_all.txt};
                \addplot[red,densely dashed,mark=none] table[x=VMs,y=N_2] {figures/vm_links_all.txt};
                \legend{$N_o$, $N_1+N_2$, $N_1$, $N_2$};
		\end{axis}
	\end{tikzpicture}
  \caption{Measured average number of 64MB FSBs used by a single VM. For VC both the number of PDS and Non-PDS FSBs used are shown.}
  \label{fig:vm-links}
\end{figure}

The full snapshot availability of a VM is estimated as follows with parameters $N_1$ and
 $N_2$ for VC and $N_o$ for VO.
Given normal data replication degree $r$, PDS data replication degree $r_c$, 
the availability of a file system block is the probability that  
all of its replicas do not appear in any group of $d$ failed machines among the total of $p$ machines. 
Namely, we define it as
\[
A(r) = 1-\binom{d}{r}/ \binom{p}{r}. 
\]
Then the availability of one VM's snapshot data under VO approach is the probability that
 all its file system blocks are unaffected during the system failure:
\begin{equation}
\label{eq:VO}
%(1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_o}. 
A(r)^{N_o}. 
\end{equation}

For VC, there are two cases for $d$ failed machines.
\begin{itemize}
\item
When $r \le d<r_c$,  there is no PDS data loss and  
the full snapshot availability of a VM in the VC approach is 
\begin{equation}
\label{eq:VC1}
%(1-\frac{\binom{d}{r}} { \binom{p}{r} })^{N_1}.
A(r)^{N_1}.
\end{equation}
Since $N_1$ is typically much smaller than $N_o$, 
the VC approach has a higher availability of VM snapshots than VO.

\item
When $r_c \leq d$, both non-PDS and PDS file blocks in VC can have a loss.
The full snapshot availability of  a VM in the VC approach is
\begin{equation}
\label{eq:VC2}
% (1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
% *
% (1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2}.
A(r)^{N_1} * A(r_c)^{N_2}
\end{equation}
Figure~\ref{fig:fsb-availability} illustrates that the availability of an individual file system block
with different replication degree. This demonstrates inequality 
% $1-\frac{ \binom{d}{r}} { \binom{p}{r} } < 1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} }$
$A(r) < A(r_c)$
.  Since $N_1$ is much smaller than $N_o$ as discussed previously, 
these two conditions support that VC has a  higher availability. 
%comparing  Formula~\ref{eq:VO} and~\ref{eq:VC}. 
\end{itemize}

\begin{figure}[htbp]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
                width=\linewidth,
                height=0.6\linewidth,
            %title={FSB Availability},
            cycle list name=mline,
            xlabel={Number of Machines Failed},
            ylabel={Availability of Single FSB (\%)},
            %extra y ticks={99.9}, %to add extra ticks
            mark options=solid,
            legend pos=south west,
            %legend columns=2,
            %legend style={
            %    at={(0.5,-0.2)},
            %anchor=north}
            ]
            \addplot table[x=NodesFailed,y=Availability5] {figures/fsb_availability.txt};
            \addplot table[x=NodesFailed,y=Availability4] {figures/fsb_availability.txt};
            \addplot table[x=NodesFailed,y=Availability3] {figures/fsb_availability.txt};
            \legend{$R=5$,$R=4$,$R=3$};
            \end{axis}
    \end{tikzpicture}
    \caption{Availability of a file system block in a 100 machine cluster with different replication 
and failure settings.THIS PLOT WILL BE REPLACED BY TABLE}
  \label{fig:fsb-availability}
\end{figure}