### 2013-09-23 snapshot_system/fast14_abstract.tex
revisions

\begin{abstract}
A cloud environment that hosts a large number of virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots. 
Deduplication of data blocks is necessary to eliminate
redundant blocks and reduce cost. Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
however, performng global deduplication and letting a data block be shared
by many virtual machines is resource expensive and less fault-resilient. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine using similarity-guided
search and associates  underlying file blocks with one VM for most cases.
It restricts cross-VM deduplication to popular blocks with extra replication support.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resources. 
This  paper  describes a comparative evaluation of this scheme in  accomplishing a competitive deduplication 
efficiency while sustaining a good backup throughput. 
\end{abstract}