### 2013-09-26 snapshot_system/fast14_abstract.tex

\begin{abstract}
A cloud environment that hosts a large number of virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots. 
Deduplication of data blocks is necessary to eliminate
redundant blocks and reduce cost. Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
however, performing global deduplication and letting a data block be shared
by many virtual machines is resource expensive and less fault-resilient. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine using similarity-guided
search and associates  underlying file blocks with one VM for most cases.
It restricts global deduplication to popular blocks with extra replication support.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resources. 
This  paper  describes a comparative evaluation of this scheme in  accomplishing competitive deduplication 
efficiency while sustaining a good backup throughput. 
\end{abstract}