### 2013-09-26 snapshot_system/fast14_abstract.tex
revisions

\begin{abstract}
A cloud environment that hosts a large number of virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots. 
Deduplication of data blocks is necessary to eliminate
redundant blocks and reduce cost. Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
however, performing global deduplication and letting a data block be shared
by many virtual machines is resource expensive and less fault-resilient. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine. Our approach uses similarity-guided
search and associate  underlying file blocks with one VM for most cases.
It restricts global deduplication to popular chunks with extra replication support.
Our analysis shows that  this VM centric scheme 
can  strike a balance for better fault isolation with competitive  deduplication efficiency
while using a small amount of computing resources. 
This  paper  describes a comparative evaluation of this scheme to assess  its deduplication 
efficiency and backup throughput. 
\end{abstract}