### 2013-10-30 thesis/Dissertation.tex
thesis template

% Root File for a UCSB Dissertation
%
%

\documentclass[12pt,twoside,draft]{ucthesis}
% todo: final options: [12pt,oneside,final]

\usepackage{graphics,epsfig}
%\usepackage{amssymb, amsmath, amsfonts, epsfig, graphicx, graphics, latexsym, makeidx, psfig, subeqn, verbatim, xspace,textcomp}

%% Document Parts

% run latex with only these files; it's faster without all eps figures etc.
%\includeonly{Title,Front,Intro}
%\includeonly{Title,Front,Intro,Literature,Chapter1,Chapter2,Conclusions}

%%% Document Portion:

\begin{document}

%% Front Matter:
\input{Title}          % First Page
\maketitle             % Make Title Page

\begin{frontmatter}
\approvalpage
\copyrightpage
\input{Dedicate}
\input{Acknowledgements}
\input{Vitae}
%\input{Preface}
\input{Abstract}
\tableofcontents
\listoffigures
\listoftables
\end{frontmatter}

% the chapters

% spacing in figures and tables and their captions can be
% changed here (\ssp for single-space, empty for same as surrounding 
% text); for this to work, the command \figsp has to be included
% in every figure and table right after the \begin{figure}
\def\figsp{\ssp}
%\def\figsp{}

\include{Intro}

%\include{Literature}

%\include{Chapter1}

%\include{Chapter2}

%\include{Conclusions}

\ssp   % bibliography can be single-spaced for UC thesis format
\bibliography{mybibfile}
\bibliographystyle{abbrv}%{unsrt}%{alpha}{siam}{apalike}{ieeetr}{plain}{acm}{abbrv}
%\bibliographystyle{IEEE-Alpha}

\part*{\addcontentsline{toc}{part}{Appendices}Appendices}
\appendix

\include{Appendix}

\end{document}