### 2013-12-10 thesis/Acknowledgements.tex

\begin{acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
I want to take this opportunity to thank my advisor Professor Tao Yang for 
his invaluable advice and instructions to me on selecting interesting and challenging
research topics, identifying specific problems for each research phase, 
as well as preparing for my future career.

I also want to thank members and ex-members of my research group, Michael Agun, and 
Hong Tang, for their incessant support in the forms of technical discussion, 
system co-development, cluster administration, and other research activities.

Finally, I want to thank my dissertation committee members, Professor 
Jianwen Su, Professor Richard Wolski 
for their insightful comments over my dissertation.

This dissertation study was supported in part by NSF xxx.
\end{acknowledgements}