### 2013-12-10 thesis/Dissertation.tex

% Root File for a UCSB Dissertation
%
%
% draft options
%\documentclass[12pt,twoside,draft]{ucthesis}
% final options: 
\documentclass[12pt,oneside,final]{ucthesis}

\usepackage{graphics,epsfig}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{sansmath}
\usepackage{mathptmx}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{breqn}
\usepackage{rotating}
%\usepackage{amssymb, amsmath, amsfonts, epsfig, graphicx, graphics, latexsym, makeidx, psfig, subeqn, verbatim, xspace,textco
%mp}
\pgfplotsset{
    %You are supposed to use append style rather than just setting
    %parameters globally so they are easier to override to special cases
    %thickness options:thin,ultra thin,very thin,semithick,thick,very thick
    every axis/.append style={
        thick,
        tick style={thick},
        tick label style={font=\small\mathversion{sans}\sffamily},
        label style={font=\small\sffamily},
        font={\small\sffamily},
        mark size=1.0
    },
    every legend/.append style={font={\small\mathversion{sans}\sffamily}}
}
\pgfplotsset{
    every mark/.append style={solid}
}
\pgfplotscreateplotcyclelist{mcolor}{%
    {blue,every mark/.append style={fill=blue},mark=*},
    {red,every mark/.append style={fill=red},mark=square*},
    {brown,every mark/.append style={fill=brown},mark=triangle*,mark size=1.5},
    {black,every mark/.append style={fill=black},mark=diamond*,mark size=1.5},
    {green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=1.4},
    {magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=1.5},
    {densely dashed,blue,every mark/.append style={fill=blue},mark=*},
    {densely dashed,red,every mark/.append style={fill=red},mark=square*},
    {densely dashed,brown,every mark/.append style={fill=brown},mark=triangle*,mark size=1.5},
    {densely dashed,black,every mark/.append style={fill=black},mark=diamond*,mark size=1.5},
    {densely dashed,green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=1.4},
    {densely dashed,magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=1.5}}

\pgfplotscreateplotcyclelist{mmark*}{
    {mark=*},
    {mark=square*},
    {mark=triangle*,mark size=1.5},
    {mark=diamond*,mark size=1.5},
    {mark=pentagon*,mark size=1.4},
    {mark=star,mark size=1.5},
    {mark=asterisk,mark size=1.5},
    {mark=+,mark size=1.5},
    {mark=Mercedes star,mark size=1.5}}

\pgfplotscreateplotcyclelist{mline}{
    {solid,blue,every mark/.append style={fill=blue}},
    {densely dashed,red,every mark/.append style={fill=red}},
    {densely dotted,brown,every mark/.append style={fill=brown}},
    {loosely dotted,black,every mark/.append style={fill=black}},
    {loosely dashed,green!60!black,every mark/.append style={fill=green!60!black}},
    {densely dashdotted,magenta!90!black,every mark/.append style={fill=magenta!90!black}},
    {densely dashdotdotted,teal,every mark/.append style={fill=teal}},
    {loosely dashdotted,orange,every mark/.append style={fill=orange}},
    {loosely dashdotdotted,violet,every mark/.append style={fill=violet}}}

    \pgfplotsset{cycle list name=mcolor}

\newcommand{\comments}[1]{}
%% Document Parts

% run latex with only these files; it's faster without all eps figures etc.
%\includeonly{Title,Front,Intro}
%\includeonly{Title,Front,Intro,Literature,Chapter1,Chapter2,Conclusions}

%%% Document Portion:

\begin{document}

%% Front Matter:
\input{Title}          % First Page
\maketitle             % Make Title Page

\begin{frontmatter}
\approvalpage
\copyrightpage
\input{Dedicate}
\input{Acknowledgements}
\input{Vitae}
%\input{Preface}
\input{Abstract}
\tableofcontents
\listoffigures
\listoftables
\end{frontmatter}

% the chapters

% spacing in figures and tables and their captions can be
% changed here (\ssp for single-space, empty for same as surrounding 
% text); for this to work, the command \figsp has to be included
% in every figure and table right after the \begin{figure}
\def\figsp{\ssp}
%\def\figsp{}

\include{Introduction}
\include{Background}
\include{Overview}
\include{OfflineDedup}
\include{InlineDedup}
\include{DataManagement}
\include{Conclusion}

\ssp   % bibliography can be single-spaced for UC thesis format
\bibliography{Dissertation}
\bibliographystyle{abbrv}%{unsrt}%{alpha}{siam}{apalike}{ieeetr}{plain}{acm}{abbrv}
%\bibliographystyle{IEEE-Alpha}

% \part*{\addcontentsline{toc}{part}{Appendices}Appendices}
% \appendix
% \include{Appendix}

\end{document}