### 2014-01-23 snapshot_system/paper.tex
add paper files

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix}
\usepackage{mathptmx}
\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{array}
\usepackage{breqn}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{sansmath}
\usepackage{booktabs}
\pgfplotsset{compat=1.5}
\DeclareGraphicsExtensions{.pdf}

\newcommand{\comments}[1]{}

%pgfplots setup
\pgfplotsset{
    %You are supposed to use append style rather than just setting
    %parameters globally so they are easier to override to special cases
    %thickness options:thin,ultra thin,very thin,semithick,thick,very thick
    every axis/.append style={
        thick,
        tick style={thick},
        tick label style={font=\small\mathversion{sans}\sffamily},
        label style={font=\small\sffamily},
        font={\small\sffamily},
        mark size=1.0
    },
    every legend/.append style={font={\small\mathversion{sans}\sffamily}}
}
\pgfplotsset{
    every mark/.append style={solid}
}
\pgfplotscreateplotcyclelist{mcolor}{%
    {blue,every mark/.append style={fill=blue},mark=*},
    {red,every mark/.append style={fill=red},mark=square*},
    {brown,every mark/.append style={fill=brown},mark=triangle*,mark size=1.5},
    {black,every mark/.append style={fill=black},mark=diamond*,mark size=1.5},
    {green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=1.4},
    {magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=1.5},
    {densely dashed,blue,every mark/.append style={fill=blue},mark=*},
    {densely dashed,red,every mark/.append style={fill=red},mark=square*},
    {densely dashed,brown,every mark/.append style={fill=brown},mark=triangle*,mark size=1.5},
    {densely dashed,black,every mark/.append style={fill=black},mark=diamond*,mark size=1.5},
    {densely dashed,green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=1.4},
    {densely dashed,magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=1.5}}

\pgfplotscreateplotcyclelist{mmark*}{
    {mark=*},
    {mark=square*},
    {mark=triangle*,mark size=1.5},
    {mark=diamond*,mark size=1.5},
    {mark=pentagon*,mark size=1.4},
    {mark=star,mark size=1.5},
    {mark=asterisk,mark size=1.5},
    {mark=+,mark size=1.5},
    {mark=Mercedes star,mark size=1.5}}

\pgfplotscreateplotcyclelist{mline}{
    {solid,blue,every mark/.append style={fill=blue}},
    {densely dashed,red,every mark/.append style={fill=red}},
    {densely dotted,brown,every mark/.append style={fill=brown}},
    {loosely dotted,black,every mark/.append style={fill=black}},
    {loosely dashed,green!60!black,every mark/.append style={fill=green!60!black}},
    {densely dashdotted,magenta!90!black,every mark/.append style={fill=magenta!90!black}},
    {densely dashdotdotted,teal,every mark/.append style={fill=teal}},
    {loosely dashdotted,orange,every mark/.append style={fill=orange}},
    {loosely dashdotdotted,violet,every mark/.append style={fill=violet}}}

    \pgfplotsset{cycle list name=mcolor}

\begin{document}

%\title{VM-Centric Deduplication with Fault Isolation for Virtual Machine Snapshot Backup }
%Collocated Snapshot Backup with VM-Centric Deduplication in the Cloud}
\title{
VM-Centric  Deduplication  with Fault Isolation for Snapshot Backup in the Cloud} 
%\author{
%  Wei Zhang$^{\star}$, Michael Agun$^\star$, Tao Yang$^\star$, and Hong Tang$^\dagger$ \\
%  {\normalsize$^\star$  University of California at Santa Barbara}, {\normalsize$^\dagger$Alibaba Inc.} \\
%  \{wei, tyang, dagun\}@cs.ucsb.edu, hongtang@alibaba-inc.com
%}

\maketitle

\input{paper_abs}
\input{paper_intro}
\input{paper_background}
\input{paper_dedup}
\input{paper_analysis}
\input{paper_del}
\input{paper_arch}
\input{paper_eval}
\input{paper_conc}

\begin{small}
\bibliographystyle{abbrv}
\bibliography{library,dedup1}
\end{small}
\end{document}