### 2014-01-28 snapshot_system/paper_intro.tex
Rich is making a pass

\section{Introduction}

Commercial ``Infrastructure as a Service'' clouds (i.e. {\em public
clouds}) often make use of commodity data center components to achieve the
best possible economies of scale.  In particular, large-scale e-commerce cloud
providers such as Google and Alibaba deploy ``converged'' components that
co-locate computing and storage in each hardware module (as opposed to having
separate computing and storage ``tiers.'')
%One of emerging architectures for building cloud services is a converged
%storage architecture that leverages commodity servers with software clustered
%storage.
%The converged architecture (server plus compute in same tier) grew out of the
%Google model. 
These providers use software to
aggregate multiple direct attached low-cost disks together across
servers as a way of avoiding the relatively high cost of network attached 
storage~\cite{googlefs03,hdfs10}.
%systems. %This(e.g. ~\cite{NutanixPaper}).  Leading cloud platform companies
%such as Google, Amazon, Alibaba, Microsoft Azure  have built a converged
%compute and storage infrastructure and used a distributed file system such as
%Google file system~\cite{googlefs03,hdfs10} to glue a large number of
%commodity servers.
% with local storage into a single cluster.
In such an environment,
%That  allows the use of scalable compute and storage, without incurring the
%costs and performance limitations associated with network storage.
each physical machine runs a number of virtual machines (VMs)
%as  instances of a guest operating system 
and their virtual disks are stored as disk image files, typically
using a local (possibly shared) file system.
Frequent snapshot backup of virtual disk images can
increase the service reliability by allowing VMs to restart from their latest
snapshot in the event of a server failure.  However, because the disk
blocks that make up the VMs are often the same from snapshot to snapshot,  
deduplication of redundant content
blocks~\cite{venti02,bottleneck08} is necessary to substantially reduce the
storage demand.
%For example, the Aliyun cloud, which is  the largest cloud service provider
%by Alibaba in China, automatically conducts  the backup of virtual disk
%images to all active users every day.  The cost of supporting a large number
%of concurrent backup streams is high because of the huge storage demand and
%the use of deduplication

%Using a separate  backup service with full deduplication support~\cite{venti02,bottleneck08}
%can effectively identify and remove content duplicates among snapshots, 
%but such a solution can be expensive. There is also a large amount of 
%network traffic to transfer  data from the host machines to the backup facility
%before duplicates are removed.

Version-based detection has been used to identify file blocks that have not
changed across snapshots so their redundant storage can be
avoided~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
%the previous version of the
%snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}, 
Alternatively, techniques that compare block ``fingerprints'' to
identify duplicates
that exist among all files regardless of their 
origin~\cite{Guo2011,Dong2011,extreme_binning09} have
also gained in popularity.

One consequence of aggressive deduplication is the possible loss of failure
resilience.  Separate files share the same physical copies of blocks that are
logically duplicated among them.  If a shared block is lost, all files that
share that block are affected.  The problem is particularly accute for VM
snapshots where the purpose of the snapshot is precisely to create redundant
data copies so that VMs are more fault resilient.  Thus, there exists a
tension between the additional redundancy that snapshot creation is intended
to provide, and the storage efficiencies that deduplication makes possible.

%Because
%of highly repetitive content in snapshots from different VMs, many data chunks
%are shared by virtual machines.  Failure of a few shared data chunks can have
%a broad effect and snapshots of these virtual machines could be affected.  The
%previous work in deduplication focuses on the efficiency and approximation of
%fingerprint comparison, and has not addressed fault tolerance issues  together
%with deduplication.

In this paper {\em we explore a method for deduplicating VM snapshots in
converged architecture IaaS systems that strikes a balance between fault
isolation and storage efficiency}.  The main contribution of this work is to
evaluate the impact of sharing only ``popular'' data blocks 
(through deduplication and limited replication)
across 
virtual machines while using localized deduplication within each VM
to achieve both storage savings and fault isolation.
%, and to propose a VM-centric approach that localizes
%deduplication as much as possible and restricts global deduplication only to a
%limited set of most popular blocks.  
%In our method, local deduplication uses
%similarity-guided elimination to improve the deduplication coverage.  
Since
the file system block size is normally bigger than the average data chunk size
used for deduplication, 
our approach uses a deduplication data chunk size that is smaller than a
single file block and
packages data chunks from the same VM into a file
system block as much as possible to improve fault isolation.  Deduplicating
data locally within a VM also reduces the need to maintain a global index of
chunks that are shared among VMs, thereby increasing the scaling efficiency of
the snapshot system.  The cost associated with localization, however, is in 
the lost
storage efficiency that cross-VM deduplication makes possible.  To recover
some of this efficiency we allow a small number of the most popular data
blocks to be deduplicated (and then replicated) across VMs.

This work explores the storage and
performance efficiencies that such a {\em VM-centric} 
combination of local and global
techniques can achieve.  We term it VM-centric because the deduplication
algorithm considers VM boundaries in making its decisions as opposed to
treating all blocks as being equivalent within the storage system.
We detail the storage, performance,
and fault isolation properties
of our method and verify these properties empirically using a prototype
system that we have developed to run as a co-located back-up service in
converged IaaS cloud enviornments.

%Because data
%sharing is restricted, this VM-centric approach reduces the overall resource
%usage significantly during backup and simplifies the snapshot deletion
%process. This low-resource design is suitable when the backup service with
%deduplication is collocated with other services running on a shared compute
%and storage cluster.  We have evaluated this VM-centric approach using  a
%prototype system.
% that runs a cluster of Linux machines with Xen and a standard distributed
% file system for the backup storage. 
%This approach localizes duplicate detection within each VM  By narrowing
%duplicate sharing within a small percent of common data chunks and exploiting
%their popularity, this scheme can afford to allocate extra replicas of these
%shared chunks for better fault resilience while sustaining competitive
%deduplication efficiency.

%In addition, our VM-centric design allows garbage collection to be performed in a localized
%scope and we propose an approximate deletion scheme to reduce this cost further.
%Localization also brings the benefits of greater ability to exploit parallelism so
%backup operations can run simultaneously without a central  bottleneck.
%This VM-centric solution uses a small amount of  memory while delivering reasonable deduplication efficiency.

%Another issue considered is that
%that garbage collection after deletion of old snapshots also competes for computing resources. 
%Sharing of data chunks among by multiple VMs needs to be detected during
%garbage collection and such dependencies complicate deletion operations.

%************** Paper sections summary
%THIS NEEDS MODIFICATION
The rest of this paper is organized as follows.
Section~\ref{sect:background} reviews the background and discusses the  design options for snapshot backup 
with a VM-centric approach. 
Section~\ref{sect:deduplication}  analyzes the tradeoff and benefits of the VM-centric approach. 
Section~\ref{sect:architecture}  describes a system implementation that evaluates the proposed techniques.
%   the benefit of our approach for fault isolation. 
Section~\ref{sect:evaluation} is our experimental evaluation that compares with other approaches.
Section~\ref{sect:conclusion}  concludes this paper.