### 2014-01-31 snapshot_system/paper_conc.tex
rev

\section{Conclusion}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is a low-profile and VM-centric deduplication scheme to 
maximize fault isolation while delivering competitive deduplication efficiency using
a small amount of system resource.
Evaluation using  this scheme strikes a tradeoff and restricted cross-VM duplicate detection
can accomplish up to 96.58\% of what complete global
deduplication can do, and the simple and approximate snapshot deletion 
effectively manages deleted chunks with a 
much lower resource usage.  The availability of snapshots increases substantially when
adding more replication for popular cross-VM chunks and packaging chunks from the same 
VM in one file system block. The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3.
%Currently we are studying how the

%Similarity guided local search reduces cross-VM data dependency and exposes more parallelism  
%while global deduplication with a small common data set eliminates popular duplicates.
%VM-specific file block packing also enhances fault tolerance by reducing data dependencies.
%The design places a special consideration for low-resource usages as a collocated cloud service
%and