### 2014-07-18 snapshot_system/paper_abs.tex
re

\begin{abstract}
Data deduplication has been widely used for cloud data backup
because of excessive redundant content. Common techniques 
perform fingerprint comparison to remove duplicates
across virtual machines.
However, fingerprint search for source-side duplicate detection is resource
intensive when the backup service is co-located with  other cloud services.
%letting a duplicate data block be shared
%by many virtual machines represents a tradeoff between
%storage efficiency and data availability.
%by many virtual machines creates data dependence and is less fault-resilient. 
%%This paper studies the impact of deduplication on 
%%fault resilidence  of virtual machine snapshots, created by data dependence  among duplicates
%%and proposes a VM-centric approach that strikes a balance between deduplication efficiency
%%and fault isolation 
This paper proposes a VM-centric backup service 
which  strikes a tradeoff for a competitive  deduplication efficiency while
has  an advantage of   using less  computing resources, suitable for running
on a converged cloud architecture that cohosts  many other services.
This VM-centric scheme also leads to  better fault isolation and simplifies snapshot deletion.  
The key techniques used are to
localize deduplication as much as possible within each 
virtual machine, guided by similarity
search, and to restrict global deduplication under popular chunks with extra replication support.
This scheme associates  underlying file blocks with one VM for most cases
and proposes an approximate method for fast and simplified snapshot deletion.
This  paper  describes an evaluation of this scheme to assess  its deduplication 
efficiency, resource usage,  and fault resilience.

% Collocating a cluster-based duplicate service with other cloud services
%   necessary to eliminate
%redundant blocks and reduce cost. Collocating a cluster-based duplicate service with other cloud services

%A cloud environment that hosts a large number of virtual machines (VMs) has
%a high storage demand for frequent backup of system image snapshots. 
%Deduplication of data blocks is necessary to eliminate
%redundant blocks and reduce cost. Collocating a cluster-based duplicate service with other cloud services
%reduces network traffic;
%however, 
\end{abstract}