### 2014-07-18 snapshot_system/paper_conc.tex

\section{Conclusion}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is a low-profile and VM-centric deduplication scheme to 
maximize fault isolation while delivering competitive deduplication efficiency using
a small amount of system resource.
Evaluation using  this scheme shows it strikes a tradeoff and restricted cross-VM duplicate detection
can accomplish up to 96.58\% of what complete global
deduplication can do, and the simple and approximate snapshot deletion 
effectively manages deleted chunks with
much lower resource usage.  The availability of snapshots increases substantially when
adding more replication for popular cross-VM chunks and packaging chunks from the same 
VM in one file system block. The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
%Currently we are studying how the

%Similarity guided local search reduces cross-VM data dependency and exposes more parallelism  
%while global deduplication with a small common data set eliminates popular duplicates.
%VM-specific file block packing also enhances fault tolerance by reducing data dependencies.
%The design places a special consideration for low-resource usages as a collocated cloud service
%and