### 2014-07-18 snapshot_system/paper_conc.tex
re

\section{Conclusion}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is a low-profile and VM-centric deduplication scheme that
uses  a small amount of system resource in both snapshot deduplication and deletion
while delivering competitive deduplication efficiency.
The VM-centric design also allows an optimization for  fast snapshot fault isolation.

Our evaluation with an Alibaba dataset has the following results.
\begin{itemize}
\item {\em Competitive deduplication efficiency with a tradeoff}. 
The proposed VC scheme 
can accomplish over  96\% of what complete global
deduplication can do. In comparison, the SRB and sampled index~\cite{Dong2011,extreme_binning09,Guo2011}
can accomplish 97\% to 97.86\%.
\item {\em Low resource usage in deduplication}. 
The VC scheme achieves a 85K:1 ratio between raw snapshot data and memory usage
and is much more memory-efficient than 
SRB  with 4K:1 ratio and sampled index with 20K:1.
Both VC and SRB are designed for a distributed cluster architecture while
the sampled index method is for a single-machine architecture.
\item {\em Low cost in snapshot deletion}. 
\item {\em Higher availability}. 
The availability of snapshots increases substantially when
adding more replication for popular cross-VM chunks and packaging chunks from the same 
VM in one file system block. The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
\end{itemize}
%Currently we are studying how the

%Similarity guided local search reduces cross-VM data dependency and exposes more parallelism  
%while global deduplication with a small common data set eliminates popular duplicates.
%VM-specific file block packing also enhances fault tolerance by reducing data dependencies.
%The design places a special consideration for low-resource usages as a collocated cloud service
%and