### 2014-07-18 snapshot_system/paper_conc.tex
rev

\section{Conclusion}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is a low-profile and VM-centric deduplication scheme that
uses  a small amount of system resource in both snapshot deduplication and deletion
while delivering competitive deduplication efficiency.
The VM-centric design also allows an optimization for  better fault isolation and tolerance.

Our evaluation with an Alibaba dataset has the following results.
\begin{itemize}
\item {\em Competitive deduplication efficiency with a tradeoff}. 
The proposed VC scheme 
can accomplish over  96\% of what complete global
deduplication can do. In comparison, SRB~\cite{Dong2011,extreme_binning09}
can accomplish 97.86 while 
the sampled index~\cite{Guo2011} can deliver 97\% for a different test dataset. 
\item {\em Lower resource usage in deduplication}. 
The VC scheme achieves a 85K:1 ratio between raw snapshot data and memory usage
and is much more memory-efficient than 
SRB  with 4K:1 ratio and sampled index with 20K:1.
VC with 100 machines takes 1.31 hours to complete the backup of 2,500 VMs
using 19\% of a single core and 16.7\% of IO bandwidth per machine. 
Processing time of VC is 23.9\% faster  than SRB in our tested cases
and the per-machine throughput of VC is reasonable based on the result 
from~\cite{Guo2011}. 
Noted that both VC and SRB are designed for a distributed cluster architecture while
the sampled index method is for a single-machine architecture.
\item {\em Lower cost in snapshot deletion}. 
The deletion in VC with summary vectors
takes less than a minute for snapshot deletion using 15MB memory per machine
and all machines run this operation in parallel without data dependence. 
The grouped mark-sweep method can take 43 hours using
1.2 to 3GB memory per machine in our tested cases.
The leak repair in VC takes 0.82 hours using 50MB on average while processing
some VMs may use upto 1.96GB. Such a repair is conducted infrequently (e.g. every 19.6 days) 
\item {\em Higher availability}. 
%The availability of snapshots increases substantially when
%adding more replication for popular cross-VM file blocks and packaging chunks from the same 
%VM in one file system block. 
The snapshot availability of VC is 99.66\% with 3 failed machines in a 100-machine cluster
while it is 69.54\% for  a VM oblivious approach.
The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
\end{itemize}
%Currently we are studying how the

%Similarity guided local search reduces cross-VM data dependency and exposes more parallelism  
%while global deduplication with a small common data set eliminates popular duplicates.
%VM-specific file block packing also enhances fault tolerance by reducing data dependencies.
%The design places a special consideration for low-resource usages as a collocated cloud service
%and