### 2014-08-29 thesis/Acknowledgements.tex
re

\begin{acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgments}

First of all, I want to take this opportunity to thank my advisor Professor Tao Yang for 
his invaluable advice and instructions to me on selecting interesting and challenging
research topics, identifying specific problems for each research phase, 
as well as preparing for my future career. I would also like to thank my committee
members, Professor Rich Wolski and Professor Jianwen Su, for their pertinent and helpful
instructions on my research work.

I also want to thank members and ex-members of my research group, Michael Agun, Gautham Narayanasamy,
Prakash Chandrasekaran, Xiaofei Du, and Hong Tang, 
for their incessant support in the forms of technical discussion, 
system co-development, cluster administration, and other research activities.

I owe my deepest gratitude to my parents, for their love and support, which give 
me the confidence and energy to overcome all past, current, and future difficulties.

Finally and most importantly, I would like to express my gratitude beyond 
words to my wife, Jiayin, who has been encouraging and inspiring me 
since we met in love. We went through all the difficult times together,
and depended on each other in this foreign place one Pacific Ocean away from our homeland.
Every step forward I have made is unimaginable without Jiayin's support.

This dissertation study was supported in part by NSF IIS-1118106.
Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and
do not necessarily reflect the views of Alibaba or the National Science Foundation.
\end{acknowledgements}