### 2015-03-06 snapshot_system/mass_abs.tex

\begin{abstract}
Data deduplication is important  for snapshot backup of virtual machines (VMs)
because of excessive redundant content.  Fingerprint search for source-side duplicate 
detection is resource intensive when the backup service for VMs  is co-located with  other 
cloud services.  This paper proposes a low-profile VM-centric backup service with  a tradeoff 
for a competitive deduplication efficiency while using small computing resources, 
suitable for running on a converged cloud architecture that cohosts  many other services.
We follow  a  VM-centric design that restricts global deduplication and 
separates popular chunks with non-popular ones.   
The  proposed techniques  are  similarity-guided local search 
for  improving the effectiveness of  this VM-centric deduplication
an  VM-centric file system block management to increase  the VM snapshot availability.  
%deduplication under popular chunks with extra replication support.
%and use an approximate method for fast and simplified snapshot deletion.
%by separating popular chunks and associating  file system blocks with one VM for non-popular chunks.
This  paper  describes an evaluation of this VM-centric scheme to assess  its deduplication 
efficiency, resource usage,  and fault tolerance.

% Collocating a cluster-based duplicate service with other cloud services
%   necessary to eliminate
%redundant blocks and reduce cost. Collocating a cluster-based duplicate service with other cloud services

%A cloud environment that hosts a large number of virtual machines (VMs) has
%a high storage demand for frequent backup of system image snapshots. 
%Deduplication of data blocks is necessary to eliminate
%redundant blocks and reduce cost. Collocating a cluster-based duplicate service with other cloud services
%reduces network traffic;
%however, 
\end{abstract}