### 2015-05-22 snapshot_system/mass.tex
add copyright info

\documentclass[conference]{IEEEtran}

\usepackage{mathptmx}
%\usepackage[colorlinks,linkcolor=red,anchorcolor=blue,citecolor=green,urlcolor=black]{hyperref}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{array}
\usepackage{breqn}
\usepackage{tikz}
\usetikzlibrary{calc,shapes,positioning,decorations.pathmorphing}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{sansmath}
\usepackage{booktabs}
\usepackage{soul}
\usepackage{flushend} %automatically balance columns of last page
\pgfplotsset{compat=1.5}
\DeclareGraphicsExtensions{.pdf}

\newcommand{\comments}[1]{}
\newcommand{\todo}[1]{\textit{\color{red}#1}}

%pgfplots setup
\pgfplotsset{
    %You are supposed to use append style rather than just setting
    %parameters globally so they are easier to override to special cases
    %thickness options:thin,ultra thin,very thin,semithick,thick,very thick
    every axis/.append style={
        thick,
        tick style={thick},
        tick label style={font=\small\mathversion{sans}\sffamily},
        label style={font=\small\sffamily},
        font={\small\sffamily},
        mark size=1.0
    },
    every legend/.append style={font={\small\mathversion{sans}\sffamily}}
}
\pgfplotsset{
    every mark/.append style={solid}
}
\pgfplotscreateplotcyclelist{mcolor}{%
    {blue,every mark/.append style={fill=blue},mark=*},
    {red,every mark/.append style={fill=red},mark=square*},
    {brown,every mark/.append style={fill=brown},mark=triangle*,mark size=1.5},
    {black,every mark/.append style={fill=black},mark=diamond*,mark size=1.5},
    {green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=1.4},
    {magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=1.5},
    {densely dashed,blue,every mark/.append style={fill=blue},mark=*},
    {densely dashed,red,every mark/.append style={fill=red},mark=square*},
    {densely dashed,brown,every mark/.append style={fill=brown},mark=triangle*,mark size=1.5},
    {densely dashed,black,every mark/.append style={fill=black},mark=diamond*,mark size=1.5},
    {densely dashed,green!60!black,every mark/.append style={fill=green!60!black},mark=pentagon*,mark size=1.4},
    {densely dashed,magenta!90!black,every mark/.append style={fill=magenta!90!black},mark=star,mark size=1.5}}

\pgfplotscreateplotcyclelist{mmark*}{
    {mark=*},
    {mark=square*},
    {mark=triangle*,mark size=1.5},
    {mark=diamond*,mark size=1.5},
    {mark=pentagon*,mark size=1.4},
    {mark=star,mark size=1.5},
    {mark=asterisk,mark size=1.5},
    {mark=+,mark size=1.5},
    {mark=Mercedes star,mark size=1.5}}

\pgfplotscreateplotcyclelist{mline}{
    {solid,blue,every mark/.append style={fill=blue}},
    {densely dashed,red,every mark/.append style={fill=red}},
    {densely dotted,brown,every mark/.append style={fill=brown}},
    {loosely dotted,black,every mark/.append style={fill=black}},
    {loosely dashed,green!60!black,every mark/.append style={fill=green!60!black}},
    {densely dashdotted,magenta!90!black,every mark/.append style={fill=magenta!90!black}},
    {densely dashdotdotted,teal,every mark/.append style={fill=teal}},
    {loosely dashdotted,orange,every mark/.append style={fill=orange}},
    {loosely dashdotdotted,violet,every mark/.append style={fill=violet}}}

    \pgfplotsset{cycle list name=mcolor}

%\linespread{.98} % to squeeeze lines and put figures in same page
\begin{document}
\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}
%\conferenceinfo{Submission to SoCC '14}{October, 2014, Seattle, WA, USA}
%\copyrightyear{2014}
%\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
%\doi{nnnnnnn.nnnnnnn}
%\title{VM-Centric Deduplication with Fault Isolation for Virtual Machine Snapshot Backup }
%Collocated Snapshot Backup with VM-Centric Deduplication in the Cloud}

%with Fault Isolation 
\title{VM-Centric Snapshot Deduplication\\ for Cloud Data Backup}
%VM-Centric  Deduplication  with Fault Isolation for Snapshot Backup in the Cloud
\comments{
\authorinfo{Wei Zhang\and Daniel Agun\and Tao Yang}
           {UC Santa Barbara}
           {\{wei,dagun,tyang\}@cs.ucsb.edu}
\authorinfo{Hong Tang}
           {Alibaba Inc.}
           {hongtang@alibaba-inc.com}
\authorinfo{Rich Wolski}
           {UC Santa Barbara}
           {rich@cs.ucsb.edu}
}

\comments{
 \authorinfo{
   Wei Zhang$^{\star\ast}$, Daniel Agun$^\star$, Tao Yang$^\star$, Rich Wolski$^{\star}$, and  
Hong Tang$^\dagger$\\
   $^\star$University of California at Santa Barbara \ \ \  \\
   $^\ast$Pure Storage Inc.\ \ \  $^\dagger$Alibaba Inc. }
%  \{wei, dagun,tyang,rich\}@cs.ucsb.edu, hongtang@alibaba-inc.com

  %{\normalsize $^\star$University of California at Santa Barbara} \ \ {\normalsize$^\dagger$Alibaba Inc.} \\
}

\author{
\IEEEauthorblockN{Wei Zhang\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Daniel Agun\IEEEauthorrefmark{1}, 
Tao Yang\IEEEauthorrefmark{1}, Rich Wolski\IEEEauthorrefmark{1}, 
Hong Tang\IEEEauthorrefmark{3}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}University of California at Santa Barbara}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Pure Storage Inc.}
\IEEEauthorblockA{\IEEEauthorrefmark{3}Alibaba Inc.}
\IEEEauthorblockA{Email: wei@purestorage.com, \{dagun, tyang, rich\}@cs.ucsb.edu, hongtang@alibaba-inc.com}
}

\IEEEoverridecommandlockouts
\IEEEpubid{978-1-4673-7619-8/15/\$31.00~\copyright~2015 IEEE}
\maketitle

\input{mass_abs}

%\category{}{Information Systems}{Information Storage Systems}
%\category{}{Distributed Architectures}{Cloud Computing}
%\category{D.2.8}{Information Systems}{Information Storage Systems}
%\terms{Design, Experimentation, Performance}

%\keywords
%Virtual machine images, snapshot backup and deletion, source-side deduplication, cluster-based architectures

\input{mass_intro}
\input{mass_background}
\input{mass_dedup}
\input{mass_analysis}
%\input{mass_del}
\input{mass_arch}
\input{mass_eval}
\input{mass_conc}

%\linespread{.97} % to squeeeze lines and put figures in same page
%\appendix
\appendices

\input{mass_dedupanalysis}
%\bibliographystyle{abbrvnat}
%\bibliographystyle{ieeetr}
\bibliographystyle{abbrv}
\bibliography{library,dedup1,dedup15}

\end{document}